---
title: "a10"
author: "Reuven Herzog"
date: "4/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This rmd document is meant to give us some aggregate counts about travel
across all modes. We will calculate:
1) Ridership on each transit route
2) Regional VMT for SOVs and HOVs
3) Regional PMT for walking and biking

```{r load libraries}
#allocating more RAM because my processor is weak, default should be 2G
options(java.parameters = "-Xmx4G")

library(here)
library(tidyverse)
library(tidytransit)
library(r5r)
library(stplanr)
library(sf)
library(PNWColors)
library(reshape2)
```

```{r load skims}
skims <- here("existing",
                     "data",
                     "all_skims_mode_numbers_ex.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  select(fromId, 
         toId, 
         n_SOV_HBW,
         n_HOV_HBW,
         n_transit_HBW,
         n_walk_HBW,
         n_bike_HBW,
         n_SOV_HBO,
         n_HOV_HBO,
         n_transit_HBO,
         n_walk_HBO,
         n_bike_HBO,
         n_SOV_NHB,
         n_HOV_NHB,
         n_transit_NHB,
         n_walk_NHB,
         n_bike_NHB,
         routes) %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId))

skims_alt <- here("alternative",
                     "data",
                     "all_skims_mode_numbers_alt.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  select(fromId, 
         toId, 
         n_SOV_HBW,
         n_HOV_HBW,
         n_transit_HBW,
         n_walk_HBW,
         n_bike_HBW,
         n_SOV_HBO,
         n_HOV_HBO,
         n_transit_HBO,
         n_walk_HBO,
         n_bike_HBO,
         n_SOV_NHB,
         n_HOV_NHB,
         n_transit_NHB,
         n_walk_NHB,
         n_bike_NHB,
         routes) %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId))
```

Now we need to turn our PA matrix into an OD matrix. We do this by averaging the
PA matrix with its transpose.

```{r PA to OD}
HBW_PA_mat <- skims %>%
  od_to_odmatrix(attrib = "n_transit_HBW", 
                 name_orig = "fromId",
                 name_dest = "toId") 

HBW_PA_mat <- HBW_PA_mat[,row.names(HBW_PA_mat)]

HBW_PA_mat_trans <- t(HBW_PA_mat)

HBW_OD_mat <- (HBW_PA_mat + HBW_PA_mat_trans) / 2

HBW_OD_table <- HBW_OD_mat %>%
  odmatrix_to_od() %>%
  rename(fromId = orig,
         toId = dest,
         flow_HBW = flow)

HBO_PA_mat <- skims %>%
  od_to_odmatrix(attrib = "n_transit_HBO", 
                 name_orig = "fromId",
                 name_dest = "toId") 

HBO_PA_mat <- HBO_PA_mat[,row.names(HBO_PA_mat)]

HBO_PA_mat_trans <- t(HBO_PA_mat)

HBO_OD_mat <- (HBO_PA_mat + HBO_PA_mat_trans) / 2

HBO_OD_table <- HBO_OD_mat %>%
  odmatrix_to_od() %>%
  rename(fromId = orig,
         toId = dest,
         flow_HBO = flow)

NHB_PA_mat <- skims %>%
  od_to_odmatrix(attrib = "n_transit_NHB", 
                 name_orig = "fromId",
                 name_dest = "toId") 

NHB_PA_mat <- NHB_PA_mat[,row.names(NHB_PA_mat)]

NHB_PA_mat_trans <- t(NHB_PA_mat)

NHB_OD_mat <- (NHB_PA_mat + NHB_PA_mat_trans) / 2

NHB_OD_table <- NHB_OD_mat %>%
  odmatrix_to_od() %>%
  rename(fromId = orig,
         toId = dest,
         flow_NHB = flow)

OD_table_all <- left_join(skims, HBW_OD_table) %>%
  left_join(HBO_OD_table) %>%
  left_join(NHB_OD_table) %>%
  select(-c(n_transit_HBW, n_transit_HBO, n_transit_NHB))
```
And now to aggregate trips by route

```{r count route}
route_trips <- OD_table_all %>%
  filter(((flow_HBW > 0 | flow_HBO > 0) | flow_NHB > 0) & !is.na(routes)) %>%
  mutate(route_1 = str_split_fixed(routes, "\\|", 3)[,1],
         route_2 = str_split_fixed(routes, "\\|", 3)[,2],
         route_3 = str_split_fixed(routes, "\\|", 3)[,3]) %>%
  pivot_longer(cols = c(route_1, route_2, route_3),
               values_to = "route") %>%
  filter(route != "") %>%
  select(route, flow_HBW, flow_HBO, flow_NHB)%>%
  group_by(route) %>%
  summarize(ridership_HBW = round(sum(flow_HBW)),
            ridership_HBO = round(sum(flow_HBO)),
            ridership_NHB = round(sum(flow_NHB))) %>% 
  mutate(total_ridership = ridership_HBW + ridership_HBO + ridership_NHB)
```


Allllrighty, let's do that whole thing for the alternative condition.

```{r PA to OD alt}
HBW_PA_mat_alt <- skims_alt %>%
  od_to_odmatrix(attrib = "n_transit_HBW", 
                 name_orig = "fromId",
                 name_dest = "toId") 

HBW_PA_mat_alt <- HBW_PA_mat_alt[,row.names(HBW_PA_mat_alt)]

HBW_PA_mat_trans_alt <- t(HBW_PA_mat_alt)

HBW_OD_mat_alt <- (HBW_PA_mat_alt + HBW_PA_mat_trans_alt) / 2

HBW_OD_table_alt <- HBW_OD_mat_alt %>%
  odmatrix_to_od() %>%
  rename(fromId = orig,
         toId = dest,
         flow_HBW = flow)


HBO_PA_mat_alt <- skims_alt %>%
  od_to_odmatrix(attrib = "n_transit_HBO", 
                 name_orig = "fromId",
                 name_dest = "toId") 

HBO_PA_mat_alt <- HBO_PA_mat_alt[,row.names(HBO_PA_mat_alt)]

HBO_PA_mat_trans_alt <- t(HBO_PA_mat_alt)

HBO_OD_mat_alt <- (HBO_PA_mat_alt + HBO_PA_mat_trans_alt) / 2

HBO_OD_table_alt <- HBO_OD_mat_alt %>%
  odmatrix_to_od() %>%
  rename(fromId = orig,
         toId = dest,
         flow_HBO = flow)


NHB_PA_mat_alt <- skims_alt %>%
  od_to_odmatrix(attrib = "n_transit_NHB", 
                 name_orig = "fromId",
                 name_dest = "toId") 

NHB_PA_mat_alt <- NHB_PA_mat_alt[,row.names(NHB_PA_mat_alt)]

NHB_PA_mat_trans_alt <- t(NHB_PA_mat_alt)

NHB_OD_mat_alt <- (NHB_PA_mat_alt + NHB_PA_mat_trans_alt) / 2

NHB_OD_table_alt <- NHB_OD_mat_alt %>%
  odmatrix_to_od() %>%
  rename(fromId = orig,
         toId = dest,
         flow_NHB = flow)


OD_table_all_alt <- left_join(skims_alt, HBW_OD_table_alt) %>%
  left_join(HBO_OD_table_alt) %>%
  left_join(NHB_OD_table_alt) %>%
  select(-c(n_transit_HBW, n_transit_HBO, n_transit_NHB))
```


```{r count route alt}
route_trips_alt <- OD_table_all_alt %>%
  filter(((flow_HBW > 0 | flow_HBO > 0) | flow_NHB > 0) & !is.na(routes)) %>%
  mutate(route_1 = str_split_fixed(routes, "\\|", 3)[,1],
         route_2 = str_split_fixed(routes, "\\|", 3)[,2],
         route_3 = str_split_fixed(routes, "\\|", 3)[,3]) %>%
  pivot_longer(cols = c(route_1, route_2, route_3),
               values_to = "route") %>%
  filter(route != "") %>%
  select(route, flow_HBW, flow_HBO, flow_NHB)%>%
  group_by(route) %>% 
  summarize(ridership_HBW = round(sum(flow_HBW)),
            ridership_HBO = round(sum(flow_HBO)),
            ridership_NHB = round(sum(flow_NHB))) %>%
  mutate(total_ridership = ridership_HBW + ridership_HBO + ridership_NHB)
```

Let's do a comparison


```{r comparison}
rtt_ex_comp <- route_trips %>%
  rename(ridership_HBW_ex = ridership_HBW,
         ridership_HBO_ex = ridership_HBO,
         ridership_NHB_ex = ridership_NHB,
         total_ridership_ex = total_ridership)

rtt_alt_comp <- route_trips_alt %>%
  rename(ridership_HBW_alt = ridership_HBW,
         ridership_HBO_alt = ridership_HBO,
         ridership_NHB_alt = ridership_NHB,
         total_ridership_alt = total_ridership)

ridership_comp <- left_join(rtt_ex_comp, rtt_alt_comp, by = "route") %>% 
  mutate(comp_HBW = ridership_HBW_alt - ridership_HBW_ex,
         comp_HBO = ridership_HBO_alt - ridership_HBO_ex,
         comp_NHB = ridership_NHB_alt - ridership_NHB_ex,
         comp_total = total_ridership_alt - total_ridership_ex)
```


Let's check the PA ridership stats and see if there are changes
```{r}
PA_ex_rides <- select(skims, c(fromId, toId,
                               n_transit_HBW, n_transit_HBO, n_transit_NHB))
PA_alt_rides <- select(skims_alt, c(fromId, toId,
                               n_transit_HBW, n_transit_HBO, n_transit_NHB)) %>%
  rename(n_transit_HBW_alt = n_transit_HBW,
         n_transit_HBO_alt = n_transit_HBO,
         n_transit_NHB_alt = n_transit_NHB)

comp_rider_PA <- left_join(PA_ex_rides, PA_alt_rides) %>%
  mutate(comp_HBW = n_transit_HBW_alt - n_transit_HBW,
         comp_HBO = n_transit_HBO_alt - n_transit_HBO,
         comp_NHB = n_transit_NHB_alt - n_transit_NHB,
         comp_all = comp_HBW+comp_HBO+comp_NHB)
```

Welp, our transit ridership pattern didn't change one bit. At least we know.


# VMT and PMT by mode

Calculating trip distance

```{r trip distances}
# centroids <- here("zones",
#                    "centroids.geojson") %>%
#   st_read() %>%
#   filter(!st_is_empty(.)) 
# 
# origins <- centroids %>% 
#   slice(rep(1:n(), each = n())) %>% 
#   mutate(dest_order = 
#            rep(seq(1:length(centroids$id)),
#                length(centroids$id)))
# 
# destinations <- origins %>%
#   arrange(dest_order)
# 
# r5r_core_existing <- here("existing",
#                           "networks") %>%
#   setup_r5(verbose = FALSE)
# 
# ped_dist <- detailed_itineraries(r5r_core_existing,
#                                  origins = origins,
#                                  destinations = destinations,
#                                  mode = "WALK",
#                                  verbose = FALSE)  %>%
#   mutate(miles = distance / 1609.34) %>%
#   select(fromId, toId, miles) %>%
#   st_drop_geometry()
# 
# drive_dist <- detailed_itineraries(r5r_core_existing,
#                                  origins = origins,
#                                  destinations = destinations,
#                                  mode = "CAR",
#                                  verbose = FALSE) %>%
#   mutate(miles = distance / 1609.34) %>%
#   select(fromId, toId, miles) %>%
#   st_drop_geometry()
# 
# bike_dist <- detailed_itineraries(r5r_core_existing,
#                                  origins = origins,
#                                  destinations = destinations,
#                                  mode = "BICYCLE",
#                                  verbose = FALSE) %>%
#   mutate(miles = distance / 1609.34) %>%
#   select(fromId, toId, miles) %>%
#   st_drop_geometry()
# 
# stop_r5() 
```

We'll also calculate trip distances for the alternative network.
This change only affects cars, so that's the only mode we need to recalculate.

```{r distance alt}
# r5r_core_alternative <- here("alternative",
#                              "networks") %>%
#   setup_r5(verbose = FALSE)
# 
# drive_dist_alt <- detailed_itineraries(r5r_core_alternative,
#                                  origins = origins,
#                                  destinations = destinations,
#                                  mode = "CAR",
#                                  verbose = FALSE) %>%
#   mutate(miles = distance / 1609.34) %>%
#   select(fromId, toId, miles) %>%
#   st_drop_geometry()
# 
# stop_r5()
```

```{r}
# # saving these to Github to avoid running the whole routing algorithm again
# 
# write.csv(drive_dist, here("assignment rmd docs", "a10", "drive_dist.csv"), row.names = FALSE)
# write.csv(drive_dist_alt, here("assignment rmd docs", "a10", "drive_dist_alt.csv"), row.names = FALSE)
# write.csv(route_trips, here("assignment rmd docs", "a10", "route_trips.csv"), row.names = FALSE)
# write.csv(route_trips_alt, here("assignment rmd docs", "a10", "route_trips_alt.csv"), row.names = FALSE)
# write.csv(ped_dist, here("assignment rmd docs", "a10", "ped_dist.csv"), row.names = FALSE)
# write.csv(bike_dist, here("assignment rmd docs", "a10", "bike_dist.csv"), row.names = FALSE)
```

```{r}
drive_dist <- read.csv(here("assignment rmd docs", "a10", "drive_dist.csv"), colClasses = c("character", "character", "numeric"))

drive_dist_alt <- read.csv(here("assignment rmd docs", "a10", "drive_dist_alt.csv"), colClasses = c("character", "character", "numeric"))

route_trips <- read.csv(here("assignment rmd docs", "a10", "route_trips.csv"),
                        colClasses = c("character", "numeric", "numeric", "numeric", "numeric"))

route_trips_alt <- read.csv(here("assignment rmd docs", "a10", "route_trips_alt.csv"),
                            colClasses = c("character", "numeric", "numeric", "numeric", "numeric"))

ped_dist <- read.csv(here("assignment rmd docs", "a10", "ped_dist.csv"), colClasses = c("character", "character", "numeric"))

bike_dist <- read.csv(here("assignment rmd docs", "a10", "bike_dist.csv"), colClasses = c("character", "character", "numeric"))
```


Calculating PMT by Mode

```{r}
PMT_car <- skims %>%
  left_join(drive_dist, by = c("fromId", "toId")) %>%
  select(n_SOV_HBW, n_HOV_HBW,
         n_SOV_HBO, n_HOV_HBO,
         n_SOV_NHB, n_HOV_NHB,
         miles) %>%
  mutate(PMT_SOV_HBW = n_SOV_HBW * miles,
         PMT_HOV_HBW = n_HOV_HBW * miles,
         PMT_SOV_HBO = n_SOV_HBO * miles,
         PMT_HOV_HBO = n_HOV_HBO * miles,
         PMT_SOV_NHB = n_SOV_NHB * miles,
         PMT_HOV_NHB = n_HOV_NHB * miles) %>%
  summarize(total_SOV_miles_HBW = sum(PMT_SOV_HBW, na.rm = TRUE),
            total_HOV_miles_HBW = sum(PMT_HOV_HBW, na.rm = TRUE),
            total_SOV_miles_HBO = sum(PMT_SOV_HBO, na.rm = TRUE),
            total_HOV_miles_HBO = sum(PMT_HOV_HBO, na.rm = TRUE),
            total_SOV_miles_NHB = sum(PMT_SOV_NHB, na.rm = TRUE),
            total_HOV_miles_NHB = sum(PMT_HOV_NHB, na.rm = TRUE)) %>%
  mutate(total_SOV_miles_all = total_SOV_miles_HBW + total_SOV_miles_HBO +
           total_SOV_miles_NHB,
         total_HOV_miles_all = total_HOV_miles_HBW + total_HOV_miles_HBO +
           total_HOV_miles_NHB)

PMT_car_alt <- skims_alt %>%
  left_join(drive_dist_alt, by = c("fromId", "toId")) %>%
  select(n_SOV_HBW, n_HOV_HBW,
         n_SOV_HBO, n_HOV_HBO,
         n_SOV_NHB, n_HOV_NHB,
         miles) %>%
  mutate(PMT_SOV_HBW = n_SOV_HBW * miles,
         PMT_HOV_HBW = n_HOV_HBW * miles,
         PMT_SOV_HBO = n_SOV_HBO * miles,
         PMT_HOV_HBO = n_HOV_HBO * miles,
         PMT_SOV_NHB = n_SOV_NHB * miles,
         PMT_HOV_NHB = n_HOV_NHB * miles) %>%
  summarize(total_SOV_miles_HBW = sum(PMT_SOV_HBW, na.rm = TRUE),
            total_HOV_miles_HBW = sum(PMT_HOV_HBW, na.rm = TRUE),
            total_SOV_miles_HBO = sum(PMT_SOV_HBO, na.rm = TRUE),
            total_HOV_miles_HBO = sum(PMT_HOV_HBO, na.rm = TRUE),
            total_SOV_miles_NHB = sum(PMT_SOV_NHB, na.rm = TRUE),
            total_HOV_miles_NHB = sum(PMT_HOV_NHB, na.rm = TRUE)) %>%
  mutate(total_SOV_miles_all = total_SOV_miles_HBW + total_SOV_miles_HBO +
           total_SOV_miles_NHB,
         total_HOV_miles_all = total_HOV_miles_HBW + total_HOV_miles_HBO +
           total_HOV_miles_NHB)


PMT_walk <- skims %>%
  left_join(ped_dist, by = c("fromId", "toId")) %>%
  select(n_walk_HBW, n_walk_HBO, n_walk_NHB, miles) %>%
  mutate(PMT_walk_HBW = n_walk_HBW * miles,
         PMT_walk_HBO = n_walk_HBO * miles,
         PMT_walk_NHB = n_walk_NHB * miles) %>%
  summarize(total_walk_miles_HBW = sum(PMT_walk_HBW, na.rm = TRUE),
            total_walk_miles_HBO = sum(PMT_walk_HBO, na.rm = TRUE),
            total_walk_miles_NHB = sum(PMT_walk_NHB, na.rm = TRUE)) %>%
  mutate(total_walk_miles_all = total_walk_miles_HBW + total_walk_miles_HBO +
           total_walk_miles_NHB)

PMT_bike <- skims %>%
  left_join(bike_dist, by = c("fromId", "toId")) %>%
  select(n_bike_HBW, n_bike_HBO, n_bike_NHB, miles) %>%
  mutate(PMT_bike_HBW = n_bike_HBW * miles,
         PMT_bike_HBO = n_bike_HBO * miles,
         PMT_bike_NHB = n_bike_NHB * miles) %>%
  summarize(total_bike_miles_HBW = sum(PMT_bike_HBW, na.rm = TRUE),
            total_bike_miles_HBO = sum(PMT_bike_HBO, na.rm = TRUE),
            total_bike_miles_NHB = sum(PMT_bike_NHB, na.rm = TRUE)) %>%
  mutate(total_bike_miles_all = total_bike_miles_HBW + total_bike_miles_HBO +
           total_bike_miles_NHB)
```

```{r}
PMT_walk_alt <- skims_alt %>%
  left_join(ped_dist, by = c("fromId", "toId")) %>%
  select(n_walk_HBW, n_walk_HBO, n_walk_NHB, miles) %>%
  mutate(PMT_walk_HBW = n_walk_HBW * miles,
         PMT_walk_HBO = n_walk_HBO * miles,
         PMT_walk_NHB = n_walk_NHB * miles) %>%
  summarize(total_walk_miles_HBW = sum(PMT_walk_HBW, na.rm = TRUE),
            total_walk_miles_HBO = sum(PMT_walk_HBO, na.rm = TRUE),
            total_walk_miles_NHB = sum(PMT_walk_NHB, na.rm = TRUE)) %>%
  mutate(total_walk_miles_all = total_walk_miles_HBW + total_walk_miles_HBO +
           total_walk_miles_NHB)

PMT_bike_alt <- skims_alt %>%
  left_join(bike_dist, by = c("fromId", "toId")) %>%
  select(n_bike_HBW, n_bike_HBO, n_bike_NHB, miles) %>%
  mutate(PMT_bike_HBW = n_bike_HBW * miles,
         PMT_bike_HBO = n_bike_HBO * miles,
         PMT_bike_NHB = n_bike_NHB * miles) %>%
  summarize(total_bike_miles_HBW = sum(PMT_bike_HBW, na.rm = TRUE),
            total_bike_miles_HBO = sum(PMT_bike_HBO, na.rm = TRUE),
            total_bike_miles_NHB = sum(PMT_bike_NHB, na.rm = TRUE)) %>%
  mutate(total_bike_miles_all = total_bike_miles_HBW + total_bike_miles_HBO +
           total_bike_miles_NHB)
```


And to calculate VMT
Using estimates for carpool sizes based on the NCHRP 716, Table 4-16

```{r}
VMT_HBW <- PMT_car$total_SOV_miles_HBW[1] +
  PMT_car$total_HOV_miles_HBW[1] / 2.42
VMT_HBO <- PMT_car$total_SOV_miles_HBO[1] +
  PMT_car$total_HOV_miles_HBO[1] / 2.71
VMT_NHB <- PMT_car$total_SOV_miles_NHB[1] +
  PMT_car$total_HOV_miles_NHB[1] / 2.75

VMT_table <- data.frame("VMT_HBW" = VMT_HBW, "VMT_HBO" = VMT_HBO,
                        "VMT_NHB" = VMT_NHB) %>%
  mutate(VMT_total = VMT_HBW + VMT_HBO + VMT_NHB)

VMT_HBW_alt <- PMT_car_alt$total_SOV_miles_HBW[1] +
  PMT_car_alt$total_HOV_miles_HBW[1] / 2.42
VMT_HBO_alt <- PMT_car_alt$total_SOV_miles_HBO[1] +
  PMT_car_alt$total_HOV_miles_HBO[1] / 2.71
VMT_NHB_alt <- PMT_car_alt$total_SOV_miles_NHB[1] +
  PMT_car_alt$total_HOV_miles_NHB[1] / 2.75

VMT_table_alt <- data.frame("VMT_HBW_alt" = VMT_HBW_alt, "VMT_HBO_alt" = VMT_HBO_alt,
                        "VMT_NHB_alt" = VMT_NHB_alt) %>%
  mutate(VMT_total_alt = VMT_HBW_alt + VMT_HBO_alt + VMT_NHB_alt)
```


```{r}
zones <- st_read(here("zones", "boundaries.geojson"))

dt_transit_ch <- comp_rider_PA %>%
  filter(toId == "48453000700"|toId == "48453001100") %>%
  nest_by(fromId) %>%
  mutate(transit_HBW_ch = sum(data$comp_HBW),
         transit_HBO_ch = sum(data$comp_HBO),
         transit_NHB_ch = sum(data$comp_NHB),
         transit_all_ch = sum(data$comp_all))

zones_ch <- zones %>%
  left_join(dt_transit_ch,
            by = c("GEOID" = "fromId"))

pal1 <- pnw_palette(name = "Bay", 
                           n = 5, 
                           type = "continuous")

ggplot(zones_ch) +
  geom_sf(aes(fill = transit_all_ch),
          color = NA) +
  scale_fill_gradientn(colours = pal1) + 
  labs(fill = "Change in Transit \nTrips to Downtown") +
  theme_void()

ggsave(here("images", "a10", "downtown_rider_map.png"))
```

```{r}
capmetro_gtfs <- read_gtfs(here("existing", "networks", "gtfs.zip"))

capmetro_shapes <- shapes_as_sf(capmetro_gtfs$shapes, crs = "NAD83")

service_types <- capmetro_gtfs$calendar_dates %>%
  filter(date == "2022-04-06")

capmetro_trips <- capmetro_gtfs$trips %>%
  filter(service_id %in% service_types$service_id) %>%
  right_join(ridership_comp,
             by = c("route_id" = "route")) %>%
  distinct(route_id, .keep_all = TRUE) %>% 
  left_join(capmetro_shapes)

capmetro_trips <- st_as_sf(capmetro_trips, sf_column_name = "geometry")
```

```{r}
bbox_new <- st_bbox(capmetro_trips) # current bounding box

xrange <- bbox_new$xmax - bbox_new$xmin # range of x values
yrange <- bbox_new$ymax - bbox_new$ymin # range of y values

scale <- .05 # percentage for padding around bbox limits

bbox_new[1] <- bbox_new[1] - (scale * xrange) # xmin - left
bbox_new[3] <- bbox_new[3] + (scale * xrange) # xmax - right
bbox_new[2] <- bbox_new[2] - (scale * yrange) # ymin - bottom
bbox_new[4] <- bbox_new[4] + (scale * yrange) # ymax - top

bbox_new <- bbox_new %>%  # take the bounding box ...
  st_as_sfc() # ... and make it a sf polygon
  
```

```{r}
ggplot() +
  geom_sf(data = capmetro_trips,
          aes(alpha = comp_total,
              geometry = geometry)) +
  scale_alpha_continuous(range = c(0.1, 1)) +
  geom_sf(data = bbox_new,
          fill = NA) +
  labs(alpha = "Increase in \nRidership") +
  theme_void()

ggsave(here("images", "a10", "route_rider_map.png"))
```
```{r}
msa_sf <- zones %>%
  st_union()

box_labx <- (st_coordinates(bbox_new)[1,1]+st_coordinates(bbox_new)[2,1])/2
  
box_laby <- (st_coordinates(bbox_new)[2,2]+st_coordinates(bbox_new)[3,2])/2

ggplot() +
  geom_sf(data = msa_sf,
          color = NA,
          fill = "gray95") + 
  geom_sf(data = capmetro_trips,
          aes(geometry = geometry)) +
  geom_sf(data = bbox_new,
          fill = "white",
          alpha = 0.9,
          size = 1) +
  geom_text(aes(x = box_labx, 
                y = box_laby,
                label = "Detail \nArea"),
            size = 8) +
  theme_void()

ggsave(here("images", "a10", "inset_map.png"))
```

```{r}
big_change <- ridership_comp %>%
  filter(comp_total > 32) %>%
  arrange(desc(comp_total))

big_change <- melt(big_change[,c("route", "total_ridership_ex", "total_ridership_alt")], id.vars = 1)
  
big_change <- big_change %>%
  mutate(variable = ifelse(variable=="total_ridership_ex","Existing","Alternative"))

ggplot(big_change,aes(x = route, y = value)) + 
    geom_bar(aes(fill = variable),stat = "identity",position = "dodge") +
  scale_fill_manual(values = c("darkorange","darkorange4")) +
  theme_minimal() + 
  labs(fill = "Condition",
       x = "Route",
       y = "Ridership")

ggsave(here("images", "a10", "bigch.png"),
        height = 4.5,
       width = 7)
```

```{r}
big_ridership <- ridership_comp %>%
  filter(total_ridership_alt > 3284) %>%
  arrange(desc(total_ridership_alt))

big_ridership <- melt(big_ridership[,c("route", "total_ridership_ex", "total_ridership_alt")], id.vars = 1)
  
big_ridership <- big_ridership %>%
  mutate(variable = ifelse(variable=="total_ridership_ex","Existing","Alternative"))

ggplot(big_ridership,aes(x = route, y = value)) + 
    geom_bar(aes(fill = variable),stat = "identity",position = "dodge") +
  scale_fill_manual(values = c("darkorange","darkorange4")) +
  theme_minimal() + 
  labs(fill = "Condition",
       x = "Route",
       y = "Ridership")

ggsave(here("images", "a10", "bigrd.png"),
       height = 4.5,
       width = 7)
```