---
title: "Assignment 8"
author: "Chris Dsida and Reuven Herzog"
date: "3/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Required Packages

This first chunk reinstalls the scenaRios package. You only need to run it once.

```{r}
library(devtools)

install_github("https://github.com/c-voulgaris/scenRios")
```

Load all the packages needed:

```{r}
library(here)
library(tidyverse)
library(sf)
library(survey)
library(srvyr)
library(od)
library(ggspatial)
library(scenRios)
```

## Loading and Manipulating Data

```{r}
zones <- here("existing",
              "data",
              "zone_data.csv") %>%
  read_csv(show_col_types = FALSE)

skims <- here("existing",
              "data",
              "all_skims_ex.csv") %>%
  read_csv(show_col_types = FALSE)
```

NHTS trip survey data is already filtered for Austin MSA:

```{r}
austin_trips <- read_csv(here(
                        "assignment rmd docs",
                        "a7",
                        "austin_trips.csv"))
```

Adding trip purposes and grouping by type:

```{r}
trips <- austin_trips %>%
  mutate(home_based = case_when(WHYTO == "01" ~ TRUE,
                                WHYTO == "02" ~ TRUE,
                                WHYFROM == "01" ~ TRUE,
                                WHYFROM == "02" ~ TRUE,
                                TRUE ~ FALSE)) %>%
  mutate(work = ifelse(WHYTO == "03" | WHYFROM == "03", TRUE, FALSE)) %>%
  mutate(purpose = case_when(home_based & work ~ "HBW",
                            home_based ~ "HBO",
                            TRUE ~ "NHB"))
```

## Summarize Average Trip Time by Purpose

Create a survey object:

```{r}
trips_svy <- trips %>%
  as_survey(weights = WTTRDFIN)

ttime_by_purpose <- trips_svy %>%
  group_by(purpose) %>%
  summarise(avg_time = survey_mean(TRVLCMIN))

ttime_by_purpose
```

## Calculate Minimum Travel Times

Finding the minimum travel time across all modes and adding it to the skims:

```{r}
skims <- skims %>%
  mutate(min_time = pmin(transit_time, 
                         car_time,
                         bike_time,
                         walk_time,
                         na.rm = TRUE))
```

## Friction Factors

A starting point for the friction factors:

```{r}
# Using Gamma Function; Large MPO 2

# skims <- skims %>%
#   mutate(F_HBW = min_time^-1.65*exp(-0.0398*min_time),
#          F_HBO = min_time^-1.51*exp(-0.18*min_time),
#          F_NHB = min_time^-1.94*exp(-0.116*min_time))
```

```{r}
# Using power function

skims <- skims %>%
  mutate(F_HBW = min_time^-2,
         F_HBO = min_time^-2,
         F_NHB = min_time^-2)
```

```{r}
# Using exp. function

# m_HBO <- ttime_by_purpose$avg_time[ttime_by_purpose$purpose == "HBO"]
# m_HBW <- ttime_by_purpose$avg_time[ttime_by_purpose$purpose == "HBW"]
# m_NHB <- ttime_by_purpose$avg_time[ttime_by_purpose$purpose == "NHB"]
# 
# skims <- skims %>%
#   mutate(F_HBO = exp(-1 * m_HBO * min_time),
#          F_HBW = exp(-1 * m_HBW * min_time),
#          F_NHB = exp(-1 * m_NHB * min_time)) 
```

## Estimate Travel Flows

Home-based other:

```{r}
HBO_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "hbo_prod",
                            zone_d = "hbo_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBO",
                            tolerance = 5,
                            max_iter = 500)
```

Home-based work:

```{r}
HBW_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "hbw_prod",
                            zone_d = "hbw_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBW",
                            tolerance = 5,
                            max_iter = 500)
```

Non-home based:

```{r}
NHB_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "nhb_prod",
                            zone_d = "nhb_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_NHB",
                            tolerance = 5,
                            max_iter = 500)
```

```{r}
tail(HBW_dist$convergence)
```

```{r}
head(HBO_dist$flows)
```

```{r}
table(HBO_dist$flows$flow > 0)
```

```{r}
convergence_points <- HBW_dist$convergence %>%
  mutate(max_diff = max_o_diff + max_d_diff) %>%
  mutate(which_max = ifelse(max_o_diff > max_d_diff, 
                            "Productions",
                            "Attractions"))

ggplot(convergence_points) +
  geom_line(aes(x = iteration, y = max_diff, lty = which_max)) +
  scale_y_continuous(name = "Maximum difference from target value",
                     trans = "log", 
                     breaks = breaks <- 10^seq(1,5, by=1),
                     labels = formatC(breaks, 
                                      big.mark = ",", 
                                      format = "f", 
                                      digits = 0)) +
  scale_x_continuous(name = "Iteration",
                     breaks = breaks <- seq(0, 14000, by=2000),
                     labels = formatC(breaks, 
                                      big.mark = ",", 
                                      format = "f", 
                                      digits = 0)) +
  scale_linetype(name = "") +
  theme_minimal()
```

```{r}
HBW_flows <- HBW_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBW_flow = flow)

HBO_flows <- HBO_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBO_flow = flow)

NHB_flows <- NHB_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         NHB_flow = flow)

skims <- skims %>%
  left_join(HBW_flows) %>%
  replace_na(list(HBW_flow = 0)) %>%
  left_join(HBO_flows) %>%
  replace_na(list(HBO_flow = 0)) %>%
  left_join(NHB_flows) %>%
  replace_na(list(NHB_flow = 0)) %>%
  mutate(total_hbw_time = HBW_flow * min_time,
         total_hbo_time = HBO_flow * min_time,
         total_nhb_time = NHB_flow * min_time)
```

```{r}
HBW_mean_time <- sum(skims$total_hbw_time) / sum(skims$HBW_flow)
HBO_mean_time <- sum(skims$total_hbo_time) / sum(skims$HBO_flow)
NHB_mean_time <- sum(skims$total_nhb_time) / sum(skims$NHB_flow)

HBW_mean_time
HBO_mean_time
NHB_mean_time
```