---
title: "Assignment 8"
author: "Chris Dsida and Reuven Herzog"
date: "3/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Required Packages

This first chunk re installs the scenaRios package. You only need to run it once.

```{r}
# library(devtools)
# 
# install_github("https://github.com/c-voulgaris/scenRios")
```

Load all the packages needed:

```{r}
library(here)
library(tidyverse)
library(sf)
library(survey)
library(srvyr)
library(od)
library(ggspatial)
library(scenRios)
```

# Existing conditions

This first section will fit a trip distribution model to our data for the existing conditions.

## Loading and Manipulating Data

```{r}
zones <- here("existing",
              "data",
              "zone_data.csv") %>%
  read_csv(show_col_types = FALSE)

skims <- here("existing",
              "data",
              "all_skims_ex.csv") %>%
  read_csv(show_col_types = FALSE)
```

NHTS trip survey data is already filtered for Austin MSA:

```{r}
austin_trips <- read_csv(here(
                        "assignment rmd docs",
                        "a7",
                        "austin_trips.csv"))
```

Adding trip purposes and grouping by type:

```{r}
trips <- austin_trips %>%
  mutate(home_based = case_when(WHYTO == "01" ~ TRUE,
                                WHYTO == "02" ~ TRUE,
                                WHYFROM == "01" ~ TRUE,
                                WHYFROM == "02" ~ TRUE,
                                TRUE ~ FALSE)) %>%
  mutate(work = ifelse(WHYTO == "03" | WHYFROM == "03", TRUE, FALSE)) %>%
  mutate(purpose = case_when(home_based & work ~ "HBW",
                            home_based ~ "HBO",
                            TRUE ~ "NHB"))
```

## Summarize Average Trip Time by Purpose

Create a survey object:

```{r}
trips_svy <- trips %>%
  as_survey(weights = WTTRDFIN)

ttime_by_purpose <- trips_svy %>%
  group_by(purpose) %>%
  summarise(avg_time = survey_mean(TRVLCMIN))

ttime_by_purpose
```

## Calculate Minimum Travel Times

Finding the minimum travel time across all modes and adding it to the skims:

```{r}
skims <- skims %>%
  mutate(min_time = pmin(transit_time, 
                         car_time,
                         bike_time,
                         walk_time,
                         na.rm = TRUE))
```

## Friction Factors

A starting point for the friction factors. Gamma function performs well.

Several options are available to us for calculating friction factors between origin and destination pairs. We assessed three different functional forms, attempting to match observed travel time values as closely as possible. Exponential functions were prohibitively time-consuming during computation and failed to improve prediction accuracy over power or gamma functions. While friction factors calculated with the power function did not have this problem, the tuneability of parameters in the gamma function led to its selection. This ease of calibration is noted in NCHRP 716 as the reason why many transportation modelers prefer to use it.

Starting values for the gamma function parameters came from the Large MPO models in Table 4.5 of NCHRP 716, which are appropriate for MSAs of greater than 1 million residents. Home-based work trip times were best approximated by the parameters in Large MPO Model 3, while Large MPO model 1 performed best initially for home-based other and non-home based trips. From these starting values, they were then adjusted until predicted average travel times by type aligned with observed data in the National Household Travel Survey. After many iterations, we were able to get modeled average travel times to within a tenth of a minute of observed values.


```{r}
# Using Gamma Function; tweaks to match average times

skims <- skims %>%
  mutate(F_HBW = min_time^-0.106*exp(-0.035*min_time),
         F_HBO = min_time^-3.693*exp(-0.015*min_time),
         F_NHB = min_time^-3.215*exp(-0.002*min_time))
```

```{r}
# Using power function

# skims <- skims %>%
#   mutate(F_HBW = min_time^-2,
#          F_HBO = min_time^-2,
#          F_NHB = min_time^-2)
```

```{r}
# Using exp. function

# m_HBO <- ttime_by_purpose$avg_time[ttime_by_purpose$purpose == "HBO"]
# m_HBW <- ttime_by_purpose$avg_time[ttime_by_purpose$purpose == "HBW"]
# m_NHB <- ttime_by_purpose$avg_time[ttime_by_purpose$purpose == "NHB"]
# 
# skims <- skims %>%
#   mutate(F_HBO = exp(-1 * m_HBO * min_time),
#          F_HBW = exp(-1 * m_HBW * min_time),
#          F_NHB = exp(-1 * m_NHB * min_time)) 
```

## Estimate Travel Flows

Home-based other:

```{r}
HBO_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "hbo_prod",
                            zone_d = "hbo_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBO",
                            tolerance = 5,
                            max_iter = 500)
```

Home-based work:

```{r}
HBW_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "hbw_prod",
                            zone_d = "hbw_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBW",
                            tolerance = 5,
                            max_iter = 500)
```

Non-home based:

```{r}
NHB_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "nhb_prod",
                            zone_d = "nhb_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_NHB",
                            tolerance = 5,
                            max_iter = 500)
```

For some reason, the balancing never seems to get to the tolerance, no matter how many iterations or which friction factors are used.

This plot was used to observe the convergence (or lack thereof).

```{r}
convergence_points <- NHB_dist$convergence %>%
  mutate(max_diff = max_o_diff + max_d_diff) %>%
  mutate(which_max = ifelse(max_o_diff > max_d_diff,
                            "Productions",
                            "Attractions"))

ggplot(filter(convergence_points, between(iteration, 0, 50))) +
  geom_line(aes(x = iteration, y = max_diff, lty = which_max)) +
  scale_y_continuous(name = "Maximum difference from target value",
                     trans = "log",
                     breaks = breaks <- 10^seq(1,5, by= 1),
                     labels = formatC(breaks,
                                      big.mark = ",",
                                      format = "f",
                                      digits = 0)) +
  scale_x_continuous(name = "Iteration",
                     breaks = breaks <- seq(0, 50, by=10),
                     labels = formatC(breaks,
                                      big.mark = ",",
                                      format = "f",
                                      digits = 0)) +
  annotate(geom = "text", x = 53, y = 407, label = "(407)", size = 3.5) +
  annotate(geom = "text", x = 53, y = 91, label = "(91)", size = 3.5) +
  scale_linetype(name = "") +
  theme_minimal()
```

Appending the various modeled trip flows to our skim table in the existing conditions.

```{r}
HBW_flows <- HBW_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBW_flow = flow)

HBO_flows <- HBO_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBO_flow = flow)

NHB_flows <- NHB_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         NHB_flow = flow)

skims <- skims %>%
  left_join(HBW_flows) %>%
  replace_na(list(HBW_flow = 0)) %>%
  left_join(HBO_flows) %>%
  replace_na(list(HBO_flow = 0)) %>%
  left_join(NHB_flows) %>%
  replace_na(list(NHB_flow = 0)) %>%
  mutate(total_hbw_time = HBW_flow * min_time,
         total_hbo_time = HBO_flow * min_time,
         total_nhb_time = NHB_flow * min_time)
```

Save the skim with travel flows to our repo.

```{r}
# write_csv(skims, here("existing",
#                       "data",
#                       "all_skims_ex.csv"))
```

This chunk used to confirm alignment with travel survey information.

```{r}
# HBW_mean_time <- sum(skims$total_hbw_time) / sum(skims$HBW_flow)
# HBO_mean_time <- sum(skims$total_hbo_time) / sum(skims$HBO_flow)
# NHB_mean_time <- sum(skims$total_nhb_time) / sum(skims$NHB_flow)
# 
# HBW_mean_time
# HBO_mean_time
# NHB_mean_time
```

# Alternative

## Importing our data

```{r}
zones_alt <- here("alternative",
                  "data",
                  "zone_data.csv") %>%
  read_csv(show_col_types = FALSE)

skims_alt <- here("alternative",
                  "data",
                  "all_skims_alt.csv") %>%
  read_csv(show_col_types = FALSE)
```

## Calculate Minimum Travel Times

Finding the minimum travel time across all modes and adding it to the skims for the alternative scenario:

```{r}
skims_alt <- skims_alt %>%
  mutate(min_time = pmin(transit_time, 
                         car_time,
                         bike_time,
                         walk_time,
                         na.rm = TRUE))
```

## Friction Factors

Using the same factors that were calibrated in the model, just including them in the alternative skim table now.

```{r}
# Using Gamma Function; tweaks to match average times

skims_alt <- skims_alt %>%
  mutate(F_HBW = min_time^-0.106*exp(-0.035*min_time),
         F_HBO = min_time^-3.693*exp(-0.015*min_time),
         F_NHB = min_time^-3.215*exp(-0.002*min_time))
```

## Estimate Travel Flows

Home-based other:

```{r}
HBO_dist_alt <- grvty_balancing(od_zones = zones_alt,
                            friction = skims_alt,
                            zone_id = "GEOID",
                            zone_o = "hbo_prod",
                            zone_d = "hbo_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBO",
                            tolerance = 5,
                            max_iter = 500)
```

Home-based work:

```{r}
HBW_dist_alt <- grvty_balancing(od_zones = zones_alt,
                            friction = skims_alt,
                            zone_id = "GEOID",
                            zone_o = "hbw_prod",
                            zone_d = "hbw_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBW",
                            tolerance = 5,
                            max_iter = 500)
```

Non-home based:

```{r}
NHB_dist_alt <- grvty_balancing(od_zones = zones_alt,
                            friction = skims_alt,
                            zone_id = "GEOID",
                            zone_o = "nhb_prod",
                            zone_d = "nhb_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_NHB",
                            tolerance = 5,
                            max_iter = 500)
```

Append the flows and travel time to the alternative skim data.

```{r}
HBW_flows_alt <- HBW_dist_alt$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBW_flow_alt = flow)

HBO_flows_alt <- HBO_dist_alt$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBO_flow_alt = flow)

NHB_flows_alt <- NHB_dist_alt$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         NHB_flow_alt = flow)

skims_alt <- skims_alt %>%
  left_join(HBW_flows_alt) %>%
  replace_na(list(HBW_flow_alt = 0)) %>%
  left_join(HBO_flows_alt) %>%
  replace_na(list(HBO_flow_alt = 0)) %>%
  left_join(NHB_flows_alt) %>%
  replace_na(list(NHB_flow_alt = 0)) %>%
  mutate(total_hbw_time_alt = HBW_flow_alt * min_time,
         total_hbo_time_alt = HBO_flow_alt * min_time,
         total_nhb_time_alt = NHB_flow_alt * min_time)
```

Adding flows from the existing conditions for the sake of comparison. Predicted absolute and percentage changes are calculated for each trip type.

```{r}
skims_ex <- skims %>%
  select(fromId, 
         toId,
         HBW_flow,
         HBO_flow,
         NHB_flow,
         total_hbw_time,
         total_hbo_time,
         total_nhb_time)

skims_alt <- skims_alt %>%
  left_join(skims_ex) 

skims_alt <- skims_alt %>%
  mutate(HBW_flow_ch = HBW_flow_alt-HBW_flow,
         HBO_flow_ch = HBO_flow_alt-HBO_flow,
         NHB_flow_ch = NHB_flow_alt-NHB_flow,
         HBW_flow_pct = HBW_flow_ch/HBW_flow,
         HBO_flow_pct = HBO_flow_ch/HBO_flow,
         NHB_flow_pct = NHB_flow_ch/NHB_flow)
```

Saving the alt skims to our repo.

```{r}
# write_csv(skims_alt, here("alternative",
#                           "data",
#                           "all_skims_alt.csv"))
```

This is for comparison of average trip time by type; all are slightly (>1 minute) longer than in the existing conditions.

```{r}
# HBW_mean_time_alt <- sum(skims_alt$total_hbw_time_alt) / sum(skims_alt$HBW_flow_alt)
# HBO_mean_time_alt <- sum(skims_alt$total_hbo_time_alt) / sum(skims_alt$HBO_flow_alt)
# NHB_mean_time_alt <- sum(skims_alt$total_nhb_time_alt) / sum(skims_alt$NHB_flow_alt)
# 
# HBW_mean_time_alt
# HBO_mean_time_alt
# NHB_mean_time_alt
```

# More Viz

```{r}
zone_boundaries <- here("zones",
              "boundaries.geojson") %>%
  st_read(quiet = TRUE)
```

```{r urban area}
urban_bounds_sf <- read_sf(here("zones",
                           "urban areas.geojson")) %>%
  filter(UATYP10 == "U") %>%
  st_transform(crs = st_crs(zone_boundaries))
```

```{r}
desire_lines_HBO <- od_to_sf(filter(HBO_dist$flows, flow >= 750), zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)

ggplot(desire_lines_HBO) +
  annotation_map_tile(type = "cartolight", zoom = 10, progress = "none") +
  geom_sf(aes(alpha = flow)) +
  theme_void()

ggsave(here("images", "a8", "desirelines750.png"))
```


```{r}
zone_boundaries <- here("zones",
              "boundaries.geojson") %>%
  st_read(quiet = TRUE)

desire_lines_HBW <- od_to_sf(filter(HBW_dist$flows, flow >= 70), zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)

ggplot(desire_lines_HBW) +
  annotation_map_tile(type = "cartolight", zoom = 10, progress = "none") +
  geom_sf(aes(alpha = flow)) +
  theme_void()

ggsave(here("images", "a8", "desirelinesHBW70.png"))
```

Comparing travel times

```{r}
ttime_by_purpose
```

```{r}
HBO_flows <- HBO_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBO_flow = flow)

skims <- skims %>%
  left_join(HBO_flows) %>%
  replace_na(list(HBO_flow = 0)) %>%
  mutate(total_time = HBO_flow * min_time)


HBO_mean_time <- sum(skims$total_time) / sum(skims$HBO_flow)

HBW_flows <- HBW_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBW_flow = flow)

skims <- skims %>%
  left_join(HBW_flows) %>%
  replace_na(list(HBW_flow = 0)) %>%
  mutate(total_time = HBW_flow * min_time)


HBW_mean_time <- sum(skims$total_time) / sum(skims$HBW_flow)

NHB_flows <- NHB_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         NHB_flow = flow)

skims <- skims %>%
  left_join(NHB_flows) %>%
  replace_na(list(NHB_flow = 0)) %>%
  mutate(total_time = NHB_flow * min_time)


NHB_mean_time <- sum(skims$total_time) / sum(skims$NHB_flow)

HBO_mean_time
HBW_mean_time
NHB_mean_time
```

```{r}
# ttime_by_purpose <- ttime_by_purpose %>%
#   add_column(pred_avg_time = c(HBO_mean_time, HBW_mean_time, NHB_mean_time))
```

```{r}
select(ttime_by_purpose, purpose, avg_time, pred_avg_time)
```


# identify sig changes between original and alternative

Calculate the change in flow for each O-D pair for each of the 3 trip types.
Then grab the 100 most extreme O-D changes for each trip type.

```{r HBO compare}
HBO_compare <- HBO_flows %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId)) %>%
  mutate(pairId = paste(fromId, toId))

temp <- HBO_flows_alt %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId)) %>%
  mutate(pairId = paste(fromId, toId))

HBO_compare <- HBO_compare %>%
  left_join(temp, by = "pairId") %>%
  select(c(-"fromId.y", -"toId.y")) %>%
  mutate(flow_diff = HBO_flow_alt - HBO_flow) %>%
  mutate(diff_absolute = abs(flow_diff)) %>%
  rename("fromId" = "fromId.x", "toId" = "toId.x")

HBO_most_extreme <- top_n(HBO_compare, 100, diff_absolute) %>%
  select(c(-"pairId", -"diff_absolute"))
```

```{r HBW compare}
HBW_compare <- HBW_flows %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId)) %>%
  mutate(pairId = paste(fromId, toId))

temp <- HBW_flows_alt %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId)) %>%
  mutate(pairId = paste(fromId, toId))

HBW_compare <- HBW_compare %>%
  left_join(temp, by = "pairId") %>%
  select(c(-"fromId.y", -"toId.y")) %>%
  mutate(flow_diff = HBW_flow_alt - HBW_flow) %>%
  mutate(diff_absolute = abs(flow_diff)) %>%
  rename("fromId" = "fromId.x", "toId" = "toId.x")

HBW_most_extreme <- top_n(HBW_compare, 100, diff_absolute) %>%
  select(c(-"pairId", -"diff_absolute"))
```

```{r NHB compare}
NHB_compare <- NHB_flows %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId)) %>%
  mutate(pairId = paste(fromId, toId))

temp <- NHB_flows_alt %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId)) %>%
  mutate(pairId = paste(fromId, toId))

NHB_compare <- NHB_compare %>%
  left_join(temp, by = "pairId") %>%
  select(c(-"fromId.y", -"toId.y")) %>%
  mutate(flow_diff = NHB_flow_alt - NHB_flow) %>%
  mutate(diff_absolute = abs(flow_diff)) %>%
  rename("fromId" = "fromId.x", "toId" = "toId.x")

NHB_most_extreme <- top_n(NHB_compare, 100, diff_absolute) %>%
  select(c(-"pairId", -"diff_absolute"))
```



Now to visualize these differences

```{r}
desire_lines_diff_HBO <- od_to_sf(HBO_most_extreme, zone_boundaries, silent = TRUE)

ggplot(desire_lines_diff_HBO) +
  annotation_map_tile(type = "cartolight", zoom = 10, progress = "none") +
  geom_sf(aes(color = flow_diff)) +
  scale_color_gradient2(low = "red", high = "blue", mid = "grey", midpoint = 0, trans = "pseudo_log") +
  theme_void()

ggsave(here("images","a8","HBO differences.png"))
```

```{r}
desire_lines_diff_NHB <- od_to_sf(NHB_most_extreme, zone_boundaries, silent = TRUE)

ggplot(desire_lines_diff_NHB) +
  annotation_map_tile(type = "cartolight", zoom = 10, progress = "none") +
  geom_sf(aes(color = flow_diff)) +
  scale_color_gradient2(low = "red", high = "blue", mid = "grey", midpoint = 0, trans = "pseudo_log") +
  theme_void()

ggsave(here("images","a8","NHB differences.png"))
```

```{r}
desire_lines_diff_HBW <- od_to_sf(HBW_most_extreme, zone_boundaries, silent = TRUE)

ggplot(desire_lines_diff_HBW) +
  annotation_map_tile(type = "cartolight", zoom = 10, progress = "none") +
  geom_sf(aes(color = flow_diff)) +
  scale_color_gradient2(low = "red", high = "blue", mid = "grey", midpoint = 0, trans = "pseudo_log") +
  theme_void()

ggsave(here("images","a8","HBW differences.png"))
```



# What zones have the greatest change overall?

Let's now aggregate the changes in flow for each tract.
Variables:
- From/to (2)
- Trip type (3)
- (Absolute / balancing (2))
We'll either have 6 or 12 maps to look through.

```{r HBO diff aggregate}

zones_flow_diff_HBO_from <- aggregate(list(HBO_from_diff = HBO_compare$flow_diff), by = list(GEOID = HBO_compare$fromId), FUN = sum)
zones_flow_diff_HBW_from <- aggregate(list(HBW_from_diff = HBW_compare$flow_diff), by = list(GEOID = HBW_compare$fromId), FUN = sum)
zones_flow_diff_NHB_from <- aggregate(list(NHB_from_diff = NHB_compare$flow_diff), by = list(GEOID = NHB_compare$fromId), FUN = sum)
zones_flow_diff_HBO_to <- aggregate(list(HBO_to_diff = HBO_compare$flow_diff), by = list(GEOID = HBO_compare$toId), FUN = sum)
zones_flow_diff_HBW_to <- aggregate(list(HBW_to_diff = HBW_compare$flow_diff), by = list(GEOID = HBW_compare$toId), FUN = sum)
zones_flow_diff_NHB_to <- aggregate(list(NHB_to_diff = NHB_compare$flow_diff), by = list(GEOID = NHB_compare$toId), FUN = sum)


zones_diff_absolute_HBO_from <- aggregate(list(HBO_from_diff = HBO_compare$diff_absolute), by = list(GEOID = HBO_compare$fromId), FUN = sum)
zones_diff_absolute_HBW_from <- aggregate(list(HBW_from_diff = HBW_compare$diff_absolute), by = list(GEOID = HBW_compare$fromId), FUN = sum)
zones_diff_absolute_NHB_from <- aggregate(list(NHB_from_diff = NHB_compare$diff_absolute), by = list(GEOID = NHB_compare$fromId), FUN = sum)
zones_diff_absolute_HBO_to <- aggregate(list(HBO_to_diff = HBO_compare$diff_absolute), by = list(GEOID = HBO_compare$toId), FUN = sum)
zones_diff_absolute_HBW_to <- aggregate(list(HBW_to_diff = HBW_compare$diff_absolute), by = list(GEOID = HBW_compare$toId), FUN = sum)
zones_diff_absolute_NHB_to <- aggregate(list(NHB_to_diff = NHB_compare$diff_absolute), by = list(GEOID = NHB_compare$toId), FUN = sum)

```

```{r}
zone_boundaries <- zone_boundaries %>%
  select(GEOID) %>%
  left_join(zones_flow_diff_HBO_from) %>%
  left_join(zones_flow_diff_HBW_from) %>%
  left_join(zones_flow_diff_NHB_from) %>%
  left_join(zones_flow_diff_HBO_to) %>%
  left_join(zones_flow_diff_HBW_to) %>%
  left_join(zones_flow_diff_NHB_to)

zone_boundaries_absolute <- zone_boundaries %>%
  select(GEOID) %>%
  left_join(zones_diff_absolute_HBO_from) %>%
  left_join(zones_diff_absolute_HBW_from) %>%
  left_join(zones_diff_absolute_NHB_from) %>%
  left_join(zones_diff_absolute_HBO_to) %>%
  left_join(zones_diff_absolute_HBW_to) %>%
  left_join(zones_diff_absolute_NHB_to)
```

Now to actually visualize the maps


```{r balanced aggregate maps}

ggplot(zone_boundaries) +
  geom_sf(aes(fill = HBO_from_diff), color = NA) +
  scale_fill_gradient2(low = "red", mid = "grey", high = "blue", midpoint = 0) +
  geom_sf(data = urban_bounds_sf, color = "black", fill = NA) +
  theme_void()

ggsave(here("images", "a8", "aggregate zones", "HBO_from.png"))


ggplot(zone_boundaries) +
  geom_sf(aes(fill = HBW_from_diff), color = NA) +
  scale_fill_gradient2(low = "red", mid = "grey", high = "blue", midpoint = 0) +
  geom_sf(data = urban_bounds_sf, color = "black", fill = NA) +
  theme_void()

ggsave(here("images", "a8", "aggregate zones", "HBW_from.png"))

ggplot(zone_boundaries) +
  geom_sf(aes(fill = NHB_from_diff), color = NA) +
  scale_fill_gradient2(low = "red", mid = "grey", high = "blue", midpoint = 0) +
  geom_sf(data = urban_bounds_sf, color = "black", fill = NA) +
  theme_void()

ggsave(here("images", "a8", "aggregate zones", "NHB_from.png"))


ggplot(zone_boundaries) +
  geom_sf(aes(fill = HBO_to_diff), color = NA) +
  scale_fill_gradient2(low = "red", mid = "grey", high = "blue", midpoint = 0) +
  geom_sf(data = urban_bounds_sf, color = "black", fill = NA) +
  theme_void()

ggsave(here("images", "a8", "aggregate zones", "HBO_to.png"))


ggplot(zone_boundaries) +
  geom_sf(aes(fill = HBW_to_diff), color = NA) +
  scale_fill_gradient2(low = "red", mid = "grey", high = "blue", midpoint = 0) +
  geom_sf(data = urban_bounds_sf, color = "black", fill = NA) +
  theme_void()

ggsave(here("images", "a8", "aggregate zones", "HBW_to.png"))

ggplot(zone_boundaries) +
  geom_sf(aes(fill = NHB_to_diff), color = NA) +
  scale_fill_gradient2(low = "red", mid = "grey", high = "blue", midpoint = 0) +
  geom_sf(data = urban_bounds_sf, color = "black", fill = NA) +
  theme_void()

ggsave(here("images", "a8", "aggregate zones", "NHB_to.png"))
```

```{r absolute aggregate maps}

ggplot(zone_boundaries_absolute) +
  geom_sf(aes(fill = HBO_from_diff), color = NA) +
  scale_fill_continuous(trans = "log") +
  geom_sf(data = urban_bounds_sf, color = "black", fill = NA) +
  theme_void()

ggsave(here("images", "a8", "aggregate zones", "abs_HBO_from.png"))


ggplot(zone_boundaries_absolute) +
  geom_sf(aes(fill = HBW_from_diff), color = NA) +
  scale_fill_continuous(trans = "log") +
  geom_sf(data = urban_bounds_sf, color = "black", fill = NA) +
  theme_void()

ggsave(here("images", "a8", "aggregate zones", "abs_HBW_from.png"))

ggplot(zone_boundaries_absolute) +
  geom_sf(aes(fill = NHB_from_diff), color = NA) +
  scale_fill_continuous(trans = "log") +
  geom_sf(data = urban_bounds_sf, color = "black", fill = NA) +
  theme_void()

ggsave(here("images", "a8", "aggregate zones", "abs_NHB_from.png"))


ggplot(zone_boundaries_absolute) +
  geom_sf(aes(fill = HBO_to_diff), color = NA) +
  scale_fill_continuous(trans = "log") +
  geom_sf(data = urban_bounds_sf, color = "black", fill = NA) +
  theme_void()

ggsave(here("images", "a8", "aggregate zones", "abs_HBO_to.png"))


ggplot(zone_boundaries_absolute) +
  geom_sf(aes(fill = HBW_to_diff), color = NA) +
  scale_fill_continuous(trans = "log") +
  geom_sf(data = urban_bounds_sf, color = "black", fill = NA) +
  theme_void()

ggsave(here("images", "a8", "aggregate zones", "abs_HBW_to.png"))

ggplot(zone_boundaries_absolute) +
  geom_sf(aes(fill = NHB_to_diff), color = NA) +
  scale_fill_continuous(trans = "log") +
  geom_sf(data = urban_bounds_sf, color = "black", fill = NA) +
  theme_void()

ggsave(here("images", "a8", "aggregate zones", "abs_NHB_to.png"))
```


