---
title: "Assignment 5"
author: "Reuven Herzog"
date: "2/27/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Calculate accessibility
We are calculating accesibility to jobs from each census tract.

We have 4 modes - driving, transit, biking, and walking.
2 network conditions - existing, and with downtown blocked off.

```{r load libraries}
#giving it a bit extra memory to compensate for my old and slow CPU
options(java.parameters = '-Xmx4G')

library(r5r)
library(here)
library(tidyverse)
library(sf)
library(lubridate)
```

We're loading up our data. Even though we have disaggregated employment measures,
we'll only use the total number of jobs for this stage. Since we are using jobs
as a proxy for activity, the type of activity isn't so relevant.

```{r load centroids}
jobs <- read_csv(
  here("existing", "data", "zone_data.csv"
       )
  ) %>%
  select(GEOID, total_emp) %>%
  mutate(GEOID = as.character(GEOID))

jobs <- here("existing",
             "data",
             "zone_data.csv") %>%
  read_csv() %>%
  select(GEOID, total_emp) %>%
  mutate(GEOID = as.character(GEOID)) %>%
  rename(id = GEOID)

centroids <- here("zones",
                  "centroids.geojson") %>%
  st_read() %>%
  left_join(jobs)

```

```{r set up core existing}
existing_core <- here("existing",
                      "networks") %>%
  setup_r5(verbose = FALSE)
```

WHAT DECAY FUNCTION SHOULD WE USE?
right now I'm choosing logistic, mu = 30, sigma = 9
average commute time around austin hovers around 30, but a noticeable number of
people commute 45 minutes, and some even over 60. So the only real way to represent
this is with a logistic slope with fat tails.


```{r car accessibility}
car_access_existing <- accessibility(existing_core,
                                     origins = centroids,
                                     destinations = centroids,
                                     opportunities_colname = 'total_emp',
                                     mode = "CAR",
                                     decay_function = "logistic",
                                     cutoffs = 30,
                                     decay_value = 9, 
                                     verbose = FALSE)
```


for bike commutes, 25 minutes feels reasonable, but might decay quickly after that
(25 minutes was my bike commute a couple of years ago, and it felt like a border line
of what was reasonable. Probably more experienced cyclists could do it regularly -
but that's the idea, 50% of people think that it's accessible, 50% don't.)

```{r bike accessibility}
bike_access_existing <- accessibility(existing_core,
                                     origins = centroids,
                                     destinations = centroids,
                                     opportunities_colname = 'total_emp',
                                     mode = "BICYCLE",
                                     decay_function = "logistic",
                                     cutoffs = 30,
                                     decay_value = 9,
                                     bike_speed = 16.09,
                                     verbose = FALSE)
```


for walking, I'd posit that anywhere inside a 20 minute walk is "close."
after that, it decays fairly linearly, but people still are willing to invest
time walking to jobs.

```{r walk accessibility}
walk_access_existing <- accessibility(existing_core,
                                     origins = centroids,
                                     destinations = centroids,
                                     opportunities_colname = 'total_emp',
                                     mode = "WALK",
                                     decay_function = "logistic",
                                     cutoffs = 30,
                                     decay_value = 9,
                                     walk_speed = 4.828,
                                     verbose = FALSE)
```


I want transit to match car accessibility

```{r transit accessibility}
transit_access_existing <- accessibility(existing_core,
                                     origins = centroids,
                                     destinations = centroids,
                                     opportunities_colname = 'total_emp',
                                     mode = "TRANSIT",
                                     decay_function = "logistic",
                                     cutoffs = 30,
                                     decay_value = 9, 
                                     verbose = FALSE,
                                     departure_datetime = 
                                       ymd_hm("2021-04-06 16:30"),
                                     time_window = 120)
```


```{r}
stop_r5()
```

aaaand save our work
```{r save results}
write_csv(car_access_existing, here(
  "assignment rmd docs", "a5", "csvs",
  "car_access_existing.csv")
  )

write_csv(bike_access_existing, here(
  "assignment rmd docs", "a5", "csvs",
  "bike_access_existing.csv")
  )

write_csv(walk_access_existing, here(
  "assignment rmd docs", "a5", "csvs",
  "walk_access_existing.csv")
  )

write_csv(transit_access_existing, here(
  "assignment rmd docs", "a5", "csvs",
  "transit_access_existing.csv")
  )
```



Now we do it all again for the alternative network


```{r set up core alt}
existing_core <- here("alternative",
                      "networks") %>%
  setup_r5(verbose = FALSE)
```

```{r car accessibility alt}
car_access_alt <- accessibility(existing_core,
                                     origins = centroids,
                                     destinations = centroids,
                                     opportunities_colname = 'total_emp',
                                     mode = "CAR",
                                     decay_function = "logistic",
                                     cutoffs = 30,
                                     decay_value = 9, 
                                     verbose = FALSE)
```


```{r bike accessibility alt}
bike_access_alt <- accessibility(existing_core,
                                     origins = centroids,
                                     destinations = centroids,
                                     opportunities_colname = 'total_emp',
                                     mode = "BICYCLE",
                                     decay_function = "logistic",
                                     cutoffs = 30,
                                     decay_value = 9,
                                     bike_speed = 16.09,
                                     verbose = FALSE)
```

```{r walk accessibility alt}
walk_access_alt <- accessibility(existing_core,
                                     origins = centroids,
                                     destinations = centroids,
                                     opportunities_colname = 'total_emp',
                                     mode = "WALK",
                                     decay_function = "logistic",
                                     cutoffs = 30,
                                     decay_value = 9,
                                     walk_speed = 4.828,
                                     verbose = FALSE)
```


```{r transit accessibility alt}
transit_access_alt <- accessibility(existing_core,
                                     origins = centroids,
                                     destinations = centroids,
                                     opportunities_colname = 'total_emp',
                                     mode = "TRANSIT",
                                     decay_function = "logistic",
                                     cutoffs = 30,
                                     decay_value = 9, 
                                     verbose = FALSE,
                                     departure_datetime = 
                                       ymd_hm("2021-04-06 16:30"),
                                     time_window = 120)
```


```{r}
stop_r5()
```

```{r save results}
write_csv(car_access_alt, here(
  "assignment rmd docs", "a5", "csvs",
  "car_access_alt.csv")
  )

write_csv(bike_access_alt, here(
  "assignment rmd docs", "a5", "csvs",
  "bike_access_alt.csv")
  )

write_csv(walk_access_alt, here(
  "assignment rmd docs", "a5", "csvs",
  "walk_access_alt.csv")
  )

write_csv(transit_access_alt, here(
  "assignment rmd docs", "a5", "csvs",
  "transit_access_alt.csv")
  )
```

```{r}
# for loading the accessibility calcs if you don't want to re-run the skims

bike_access_alt <- read.csv(here("assignment rmd docs", 
                                 "a5", 
                                 "csvs",
                                 "bike_access_alt.csv"))

bike_access_existing <- read.csv(here("assignment rmd docs", 
                                 "a5", 
                                 "csvs",
                                 "bike_access_existing.csv"))

car_access_alt <- read.csv(here("assignment rmd docs", 
                                 "a5", 
                                 "csvs",
                                 "car_access_alt.csv"))

car_access_existing <- read.csv(here("assignment rmd docs", 
                                 "a5", 
                                 "csvs",
                                 "car_access_existing.csv"))

transit_access_alt <- read.csv(here("assignment rmd docs", 
                                 "a5", 
                                 "csvs",
                                 "transit_access_alt.csv"))

transit_access_existing <- read.csv(here("assignment rmd docs", 
                                 "a5", 
                                 "csvs",
                                 "transit_access_existing.csv"))

walk_access_alt <- read.csv(here("assignment rmd docs", 
                                 "a5", 
                                 "csvs",
                                 "walk_access_alt.csv"))

walk_access_existing <- read.csv(here("assignment rmd docs", 
                                 "a5", 
                                 "csvs",
                                 "walk_access_existing.csv"))
```


#Comparisons

```{r car/transit ratio existing}
car_access_existing <- car_access_existing %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         car_access = accessibility)

transit_access_existing <- transit_access_existing %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         transit_access = accessibility)

bike_access_existing <- bike_access_existing %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         bike_access = accessibility)

walk_access_existing <- walk_access_existing %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         walk_access = accessibility)

access_compare <- left_join(car_access_existing, transit_access_existing) %>%
  left_join(bike_access_existing) %>%
  left_join(walk_access_existing) %>%
  mutate(transit_car_ratio = transit_access / car_access,
         bike_car_ratio = bike_access / car_access,
         walk_car_ratio = walk_access / car_access)
```

```{r car/transit ratio alt}
car_access_alt <- car_access_alt %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         car_access = accessibility)

transit_access_alt <- transit_access_alt %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         transit_access = accessibility)

bike_access_alt <- bike_access_alt %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         bike_access = accessibility)

walk_access_alt <- walk_access_alt %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         walk_access = accessibility)

access_compare_alt <- left_join(car_access_alt, transit_access_alt) %>%
  left_join(bike_access_alt) %>%
  left_join(walk_access_alt) %>%
  mutate(transit_car_ratio = transit_access / car_access,
         bike_car_ratio = bike_access / car_access,
         walk_car_ratio = walk_access / car_access)
```

```{r save compares}
write_csv(access_compare, file = here("existing", 
                                      "data",
                                      "access_compare.csv"))

write_csv(access_compare_alt, file = here("alternative", 
                                      "data",
                                      "access_compare_alt.csv"))
```


#Graphics
now we need graphics - that will be in a different document in the same folder.
