---
title: "Assignment 5 graphics"
author: "Reuven Herzog"
date: "2/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

It's just easier to keep the accessibility skims and graphics separately, so we
can update in parallel if necessary.

```{r load libraries}
library(tidyverse)
library(here)
library(sf)
library(ggthemes)
library(RColorBrewer)
library(PNWColors)
```

Now we need to pull our data from csv form
```{r load data}
car_existing <- read_csv(
  here("assignment rmd docs","a5",
  "car_access_existing.csv")
  ) %>%
  select(from_id, accessibility) %>%
  mutate(from_id = as.character(from_id)) %>%
  rename("car_access" = accessibility)

bike_existing <- read_csv(
  here("assignment rmd docs","a5",
  "bike_access_existing.csv")
  ) %>%
  select(from_id, accessibility) %>%
  mutate(from_id = as.character(from_id)) %>%
  rename("bike_access" = accessibility)

walk_existing <- read_csv(
  here("assignment rmd docs","a5",
  "walk_access_existing.csv")
  ) %>%
  select(from_id, accessibility) %>%
  mutate(from_id = as.character(from_id)) %>%
  rename("walk_access" = accessibility)

transit_existing <- read_csv(
  here("assignment rmd docs","a5",
  "transit_access_existing.csv")
  ) %>%
  select(from_id, accessibility) %>%
  mutate(from_id = as.character(from_id)) %>%
  rename("transit_access" = accessibility)

#Now the alts
car_alt <- read_csv(
  here("assignment rmd docs","a5",
  "car_access_alt.csv")
  ) %>%
  select(from_id, accessibility) %>%
  mutate(from_id = as.character(from_id)) %>%
  rename("car_access" = accessibility)

bike_alt <- read_csv(
  here("assignment rmd docs","a5",
  "bike_access_alt.csv")
  ) %>%
  select(from_id, accessibility) %>%
  mutate(from_id = as.character(from_id)) %>%
  rename("bike_access" = accessibility)

walk_alt <- read_csv(
  here("assignment rmd docs","a5",
  "walk_access_alt.csv")
  ) %>%
  select(from_id, accessibility) %>%
  mutate(from_id = as.character(from_id)) %>%
  rename("walk_access" = accessibility)

transit_alt <- read_csv(
  here("assignment rmd docs","a5",
  "transit_access_alt.csv")
  ) %>%
  select(from_id, accessibility) %>%
  mutate(from_id = as.character(from_id)) %>%
  rename("transit_access" = accessibility)

#Now the ratios
access_compare_existing = read_csv(here("existing", 
                                      "data",
                                      "access_compare.csv")) %>%
  select(GEOID, ratio) %>%
  mutate(GEOID = as.character(GEOID)) %>%
  rename("car_trans_ratio_ex" = ratio)

access_compare_alt = read_csv(here("alternative", 
                                      "data",
                                      "access_compare.csv")) %>%
  select(GEOID, ratio) %>%
  mutate(GEOID = as.character(GEOID)) %>%
  rename("car_trans_ratio_alt" = ratio)
```
```{r join tables}
all_access_existing <- left_join(car_existing, bike_existing) %>%
  left_join(walk_existing) %>%
  left_join(transit_existing) %>%
  left_join(access_compare_existing, by = c("from_id" = "GEOID"))

all_access_alt <- left_join(car_alt, bike_alt) %>%
  left_join(walk_alt) %>%
  left_join(transit_alt) %>%
  left_join(access_compare_alt, by = c("from_id" = "GEOID"))
```
Now to combine the accessibility calculations with our TAZ boundaries. The last bit of code in this chunk calculates the percentage change in car vs. transit accessibility ratios.

```{r join with spatial data}
zones <- st_read(here("zones", "boundaries.geojson"))

all_access_existing <- left_join(zones, 
                                 all_access_existing, 
                                 by = c("GEOID" = "from_id"))
  
all_access_alt <- left_join(zones,
                            all_access_alt,
                            by = c("GEOID" = "from_id")) %>%
  left_join(access_compare_existing) %>%
  mutate(ratio_change = (car_trans_ratio_alt - car_trans_ratio_ex) / car_trans_ratio_ex)
```

The following code chunks all have the same structure: define a palette (one for each mode), create map in ggplot, and save it as a png to the repo. For each mode, maps are made for the existing conditions and the alternative. Car/transit accessibility ratios are compared as well. There are no changes in job accessibility for any modes other than driving, though this does result in changes to the access ratios.

```{r car accessibility in the existing conditions}
car_pal <- pnw_palette(name = "Moth", 
                           n = 5, 
                           type = "continuous")

car_access_ex <- ggplot(all_access_existing) +
  geom_sf(aes(fill = car_access),
          color = NA) +
  scale_fill_gradientn(colours = car_pal,
                       name = "Job Accessibility \nby Driving (Existing)",
                       breaks = size_brks <- seq(0, 1000000, by = 200000),
                       labels = formatC(size_brks, format = "d", big.mark = ",")) +
  theme_void()

car_access_ex

ggsave("car_access_ex.png",
       path = here("images",
                   "a5"),
       width = 7.5,
       height = 7.5,
       units = "in")
```
```{r transit accessibility in the existing conditions}
tr_pal <- pnw_palette(name = "Starfish", 
                           n = 5, 
                           type = "continuous")

trans_access_ex <- ggplot(all_access_existing) +
  geom_sf(aes(fill = transit_access),
          color = NA) +
  scale_fill_gradientn(colours = tr_pal,
                       name = "Job Accessibility \nby Transit (Existing)",
                       breaks = size_brks <- seq(0, 125000, by = 25000),
                       labels = formatC(size_brks, format = "d", big.mark = ",")) +
  theme_void()

trans_access_ex

ggsave("trans_access_ex.png",
       path = here("images",
                   "a5"),
       width = 7.5,
       height = 7.5,
       units = "in")
```

```{r car accessibility in the alternative}
car_access_alt <- ggplot(all_access_alt) +
  geom_sf(aes(fill = car_access),
          color = NA) +
  scale_fill_gradientn(colours = car_pal,
                       name = "Job Accessibility \nby Driving (Alternative)",
                       breaks = size_brks <- seq(0, 1000000, by = 200000),
                       labels = formatC(size_brks, format = "d", big.mark = ",")) +
  theme_void()

car_access_alt

ggsave("car_access_alt.png",
       path = here("images",
                   "a5"),
       width = 7.5,
       height = 7.5,
       units = "in")
```
```{r transit accessibility in the alternative}
trans_access_alt <- ggplot(all_access_alt) +
  geom_sf(aes(fill = transit_access),
          color = NA) +
  scale_fill_gradientn(colours = tr_pal,
                       name = "Job Accessibility \nby Transit (Alternative)",
                       breaks = size_brks <- seq(0, 125000, by = 25000),
                       labels = formatC(size_brks, format = "d", big.mark = ",")) +
  theme_void()

trans_access_alt

ggsave("trans_access_alt.png",
       path = here("images",
                   "a5"),
       width = 7.5,
       height = 7.5,
       units = "in")
```
```{r bike accessibility in existing conditions}
bk_pal <- pnw_palette(name = "Sunset2", 
                           n = 5, 
                           type = "continuous")

bike_access_ex <- ggplot(all_access_existing) +
  geom_sf(aes(fill = bike_access),
          color = NA) +
  scale_fill_gradientn(colours = bk_pal,
                       name = "Job Accessibility \nby Bicycle (Existing)",
                       breaks = size_brks <- seq(0, 250000, by = 50000),
                       labels = formatC(size_brks, format = "d", big.mark = ",")) +
  theme_void()

bike_access_ex

ggsave("bike_access_ex.png",
       path = here("images",
                   "a5"),
       width = 7.5,
       height = 7.5,
       units = "in")
```
```{r bike accessibility in alternative}
bike_access_alt <- ggplot(all_access_alt) +
  geom_sf(aes(fill = bike_access),
          color = NA) +
  scale_fill_gradientn(colours = bk_pal,
                       name = "Job Accessibility \nby Bicycle (Alternative)",
                       breaks = size_brks <- seq(0, 250000, by = 50000),
                       labels = formatC(size_brks, format = "d", big.mark = ",")) +
  theme_void()

bike_access_ex

ggsave("bike_access_alt.png",
       path = here("images",
                   "a5"),
       width = 7.5,
       height = 7.5,
       units = "in")
```

```{r walking accessibility in the existing conditions}
walk_pal <- pnw_palette(name = "Bay", 
                           n = 5, 
                           type = "continuous")

walk_access_ex <- ggplot(all_access_existing) +
  geom_sf(aes(fill = walk_access),
          color = NA) +
  scale_fill_gradientn(colours = walk_pal,
                       name = "Job Accessibility \nby Walking (Existing)",
                       breaks = size_brks <- seq(0, 100000, by = 20000),
                       labels = formatC(size_brks, format = "d", big.mark = ",")) +
  theme_void()

walk_access_ex

ggsave("walk_access_ex.png",
       path = here("images",
                   "a5"),
       width = 7.5,
       height = 7.5,
       units = "in")
```
```{r walking accessibility in the alternative}
walk_access_alt <- ggplot(all_access_alt) +
  geom_sf(aes(fill = walk_access),
          color = NA) +
  scale_fill_gradientn(colours = walk_pal,
                       name = "Job Accessibility \nby Walking (Alternative)",
                       breaks = size_brks <- seq(0, 100000, by = 20000),
                       labels = formatC(size_brks, format = "d", big.mark = ",")) +
  theme_void()

walk_access_alt

ggsave("walk_access_alt.png",
       path = here("images",
                   "a5"),
       width = 7.5,
       height = 7.5,
       units = "in")
```

```{r ratio of car to transit accessibility, existing conditions}
ratio_pal <- brewer.pal(5, "YlOrBr")

ratio_ex <- ggplot(all_access_existing) +
  geom_sf(aes(fill = car_trans_ratio_ex),
          color = NA) +
  scale_fill_gradientn(colours = ratio_pal,
                       name = "Transit/Driving Access \nRatio (Existing)") +
  theme_void()

ratio_ex

ggsave("ratio_ex.png",
       path = here("images",
                   "a5"),
       width = 7.5,
       height = 7.5,
       units = "in")
```
```{r ratio of car to transit accessibility, alternative}
ratio_alt <- ggplot(all_access_alt) +
  geom_sf(aes(fill = car_trans_ratio_alt),
          color = NA) +
  scale_fill_gradientn(colours = ratio_pal,
                       name = "Transit/Driving Access \nRatio (Alternative)") +
  theme_void()

ratio_alt

ggsave("ratio_alt.png",
       path = here("images",
                   "a5"),
       width = 7.5,
       height = 7.5,
       units = "in")
```
```{r change in access ratios}
car_trans_change <- ggplot(all_access_alt) +
  geom_sf(aes(fill = ratio_change),
          color = NA) +
  scale_fill_gradientn(colours = ratio_pal,
                       name = "Change in Transit/Driving \nAccess Ratio",
                       labels = scales::percent) +
  theme_void()

car_trans_change

ggsave("ratio_change.png",
       path = here("images",
                   "a5"),
       width = 7.5,
       height = 7.5,
       units = "in")
```