---
title: "Assignment 5 graphics"
author: "Reuven Herzog"
date: "2/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

It's just easier to keep the accessibility skims and graphics separately, so we
can update in parallel if necessary.

```{r load libraries}
library(tidyverse)
library(here)
library(sf)
library(ggthemes)
library(RColorBrewer)
```

Now we need to pull our data from csv form
```{r load data}
car_existing <- read_csv(
  here("assignment rmd docs","a5",
  "car_access_existing.csv")
  ) %>%
  mutate(from_id = as.character(from_id))

bike_existing <- read_csv(
  here("assignment rmd docs","a5",
  "bike_access_existing.csv")
  ) %>%
  mutate(from_id = as.character(from_id))

walk_existing <- read_csv(
  here("assignment rmd docs","a5",
  "walk_access_existing.csv")
  ) %>%
  mutate(from_id = as.character(from_id))

transit_existing <- read_csv(
  here("assignment rmd docs","a5",
  "transit_access_existing.csv")
  ) %>%
  mutate(from_id = as.character(from_id))

#Now the alts
car_alt <- read_csv(
  here("assignment rmd docs","a5",
  "car_access_alt.csv")
  ) %>%
  mutate(from_id = as.character(from_id))

bike_alt <- read_csv(
  here("assignment rmd docs","a5",
  "bike_access_alt.csv")
  ) %>%
  mutate(from_id = as.character(from_id))

walk_alt <- read_csv(
  here("assignment rmd docs","a5",
  "walk_access_alt.csv")
  ) %>%
  mutate(from_id = as.character(from_id))

transit_alt <- read_csv(
  here("assignment rmd docs","a5",
  "transit_access_alt.csv")
  ) %>%
  mutate(from_id = as.character(from_id))

#Now the ratios
access_compare_existing = read_csv(here("existing", 
                                      "data",
                                      "access_compare.csv")) %>%
  mutate(GEOID = as.character(GEOID))

access_compare_alt = read_csv(here("alternative", 
                                      "data",
                                      "access_compare.csv")) %>%
  mutate(GEOID = as.character(GEOID))
```
Alright, here are the maps we need, at minimum:
Chloropeth maps for accessibility by tract for each mode for each network condition
(8 maps minimum)
Plus maps that show transit/car ratio
Total 10 maps

Any more maps as we think of them

Graphs?
Histograms
Transit/car ratio vs. car accessibility

Should we have an absolute scale that holds across modes?
