---
title: "Assigment 2"
author: "Chris Dsida & Reuven Herzog"
date: "2/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this R Markdown document is to assemble a dataset for building our travel model. 
First, we import ACS data with population characteristics. 
Next, we download employment information for the MSA and join it with the census data. 
Additional variables are calculated from this. 
Finally, the spatial information is separated from the data, and they are both uploaded to Github. 
The second half of this document generates data visualizations used in our report.

```{r load libraries}
# load packages

library(tidycensus)
library(tidyverse)
library(units)
library(sf)
library(knitr)
library(PNWColors)
library(lehdr)
library(here)
```

```{r}
# view ACS variables

varlist_2019acs5 <- load_variables(2019, "acs5")

# look up Austin MSA GEOID

geoidlist <- get_acs(geography = "metropolitan statistical area/micropolitan statistical area",
                     variables = "B19001_001",
                     year = 2019)
```

```{r}
# # estimate income quintiles for MSA
# 
# inc_vars <- c(inc_lt_10k = 'B19001_002',
# inc_btw_10k_15k = 'B19001_003',
# inc_btw_15k_20k = 'B19001_004',
# inc_btw_20k_25k = 'B19001_005',
# inc_btw_25k_30k = 'B19001_006',
# inc_btw_30k_35k = 'B19001_007',
# 
# inc_btw_35k_40k = 'B19001_008',
# inc_btw_40k_45k = 'B19001_009',
# inc_btw_45k_50k = 'B19001_010',
# inc_btw_50k_60k = 'B19001_011',
# 
# inc_btw_60k_75k = 'B19001_012',
# inc_btw_75k_100k = 'B19001_013',
# 
# inc_btw_100k_125k = 'B19001_014',
# inc_btw_125k_150k = 'B19001_015',
# 	
# inc_btw_150k_200k = 'B19001_016',
# inc_gt_200k = 'B19001_017')
# 
# # import income data from ACS
# 
# region_inc <- get_acs(
#   geography = "metropolitan statistical area/micropolitan statistical area",
#   variables = inc_vars,
#   summary_var = 'B19001_001',
#   geometry = FALSE) %>%
#   filter(GEOID == "12420") %>%
#   mutate(pct = estimate / summary_est) %>%
#   mutate(cumul_pct = cumsum(pct)) %>%
#   select(variable, cumul_pct)
# 
# # display cumulative percentiles
# 
# kable(region_inc, digits=2)
```

```{r}
# define variable list

vars <- c(
          hh_total = 'B08201_001',
          hh_no_veh = 'B08201_002',
          workers_total = 'B08203_001',
          workers_no_veh = 'B08203_002',
          agg_vehicles = 'B25046_001',

          housing_units = 'B25001_001',
          
          commute_time = 'B08013_001',

          hh_1person = 'B08201_007',
          hh_2person = 'B08201_013',
          hh_3person = 'B08201_019',
          hh_4person_plus = 'B08201_025',

          pop_under_18 = 'B09001_001',
          pop_over_65 = 'B08101_008',
          
          pop_total = 'B02001_001',
          pop_white = 'B02001_002',
          pop_black = 'B02001_003',
          pop_ai = 'B02001_004',
          pop_asian = 'B02001_005',
          pop_nhpi = 'B02001_006',
          pop_other = 'B02001_007',
          pop_two_plus = 'B02001_010',
          pop_hisp_lat = 'B03002_012',

          inc_lt_10k = 'B19001_002',
          inc_btw_10k_15k = 'B19001_003',
          inc_btw_15k_20k = 'B19001_004',
          inc_btw_20k_25k = 'B19001_005',
          inc_btw_25k_30k = 'B19001_006',
          inc_btw_30k_35k = 'B19001_007',
          inc_btw_35k_40k = 'B19001_008',
          inc_btw_40k_45k = 'B19001_009',
          inc_btw_45k_50k = 'B19001_010',
          inc_btw_50k_60k = 'B19001_011',
          inc_btw_60k_75k = 'B19001_012',
          inc_btw_75k_100k = 'B19001_013',
          inc_btw_100k_125k = 'B19001_014',
          inc_btw_125k_150k = 'B19001_015',
          inc_btw_150k_200k = 'B19001_016',
          inc_gt_200k = 'B19001_017')

counties <- c("Travis", 
              "Bastrop", 
              "Hays", 
              "Caldwell", 
              "Williamson")
```

```{r}
#import tract-level data

census_data <- get_acs(geography = "tract",
                       state = "TX",
                       variables = vars,
                       county = counties,
                       output = "wide",
                       geometry = TRUE)
```

```{r}
# convert income ranges to approximate quintiles

census_data <- census_data %>%
  mutate(inc_quint_1 = inc_lt_10kE +
                       inc_btw_10k_15kE +
                       inc_btw_15k_20kE +
                       inc_btw_20k_25kE +
                       inc_btw_25k_30kE+
                       inc_btw_30k_35kE,
         inc_quint_2 = inc_btw_35k_40kE +
                       inc_btw_40k_45kE +
                       inc_btw_45k_50kE +
                       inc_btw_50k_60kE,
         inc_quint_3 = inc_btw_60k_75kE +
                       inc_btw_75k_100kE,
         inc_quint_4 = inc_btw_100k_125kE +
                       inc_btw_125k_150kE,
         inc_quint_5 = inc_btw_150k_200kE +
                       inc_gt_200kE)

# calculate tract area in square miles

census_data <- census_data %>%
  mutate(tract_area = st_area(geometry) %>%
           set_units("mi2"))

# additional variable generation

census_data <- census_data %>%
  mutate(pct_low_inc = inc_quint_1/hh_totalE, 
         pct_high_inc = inc_quint_5/hh_totalE,
         res_density = housing_unitsE/tract_area,
         avg_commute_time = commute_timeE/workers_totalE,
         veh_per_hh = agg_vehiclesE/hh_totalE)

# convert race/ethnicity counts to percentages
    
census_data <- census_data %>%
  mutate(pct_white = pop_whiteE/pop_totalE,
         pct_nonwhite = 1-pct_white,
         pct_black = pop_blackE/pop_totalE,
         pct_ai = pop_aiE/pop_totalE,
         pct_asian = pop_asianE/pop_totalE,
         pct_nhpi = pop_nhpiE/pop_totalE,
         pct_other = pop_otherE/pop_totalE,
         pct_two_plus = pop_two_plusE/pop_totalE,
         pct_hisp_lat = pop_hisp_latE/pop_totalE)
```

```{r lodes data}
# pull GEOIDS for tracts in the MSA

MSA <- get_acs(geography = "tract", 
               state = "TX", 
               variables = "B19001_001",
               county = counties, 
               geometry = TRUE)

MSA_GEOIDS <- MSA$GEOID

# import employment data for tracts in list

lodes <- grab_lodes(state = "tx", 
                    year = 2019, 
                    lodes_type = "wac", 
                    job_type = "JT00", 
                    segment = "S000", 
                    state_part = "main", 
                    agg_geo = "tract") %>%
  filter(w_tract %in% MSA_GEOIDS)%>%
    rename(GEOID = w_tract) %>%
    rename(total_emp = C000) %>%
    mutate(basic_emp = CNS01 + CNS02 + CNS03 + CNS04 + 
                       CNS05 + CNS06 + CNS08 + CNS09) %>%
    rename(retail_emp = CNS07) %>%
    mutate(service_emp = total_emp - basic_emp - retail_emp) %>%
    select(GEOID, total_emp, basic_emp, retail_emp, service_emp)

commute_od <- read_csv(here("commutes",
                            "O-D commutes all MSA.csv"), 
                       col_types = "ncccnnnnnnnnnn")

# extract commute origin tracts

commute_origin <- commute_od %>%
  select(w_tract, h_tract, S000) %>%
  group_by(h_tract) %>%
  rename(GEOID = h_tract) %>%
  summarize(commutes = sum(S000))

# append commute origin to data

MSA <- left_join(MSA, commute_origin)

# create object for downtown tracts affected by alternative

downtown <- st_union(filter(MSA,GEOID == "48453001100"),
                     filter(MSA,GEOID == "48453000700"))

# define function for primary employment sector in a tract

prim_emp <- function(basic, retail, service) {
  if((basic >= retail) & (basic >= service))
    return("basic")
  else if(retail >= service)
    return("retail")
  else return("service")
}
```

```{r}
# merge LEHD and ACS data

zones <- left_join(census_data, lodes)
```

```{r}
# calculate employment statistics

zones <- zones %>%
  mutate(job_density = total_emp/tract_area,
         job_home_ratio = total_emp/housing_unitsE,
         primary_emp_type = prim_emp(basic_emp, retail_emp, service_emp))
```

```{r}
# remove unnecessary columns from dataset

zones <- zones %>%
  select(GEOID,
         total_emp,
         basic_emp,
         retail_emp,
         service_emp,
         primary_emp_type,
         job_density,
         job_home_ratio,
         workers_totalE,
         workers_no_vehE,
         avg_commute_time,
         hh_totalE,
         hh_no_vehE,
         agg_vehiclesE,
         veh_per_hh,
         tract_area,
         housing_unitsE,
         res_density,
         hh_1personE,
         hh_2personE,
         hh_3personE,
         hh_4person_plusE,
         pop_totalE,
         pop_under_18E,
         pop_over_65E,
         inc_quint_1,
         inc_quint_2,
         inc_quint_3,
         inc_quint_4,
         inc_quint_5,
         pct_low_inc,
         pct_high_inc,
         pct_white,
         pct_nonwhite,
         pct_black,
         pct_ai,
         pct_asian,
         pct_nhpi,
         pct_other,
         pct_two_plus,
         pct_hisp_lat)
```

```{r}
# separate zone geometry from data

zone_boundaries <- zones %>%
  select(GEOID, geometry)

zone_data <- zones %>%
  st_drop_geometry()

# save zone and data to repo separately as .csv files

st_write(zone_boundaries, "zones/boundaries.geojson", append = FALSE)

write_csv(zone_data, "existing/data/zone_data.csv", append = FALSE)

write_csv(zone_data, "alternative/data/zone_data.csv", append = FALSE)
```

```{r}
zone_data <- read.csv(here("existing", "data", "zone_data.csv"))
zone_data$GEOID <- as.character(zone_data$GEOID)

zones <- st_read(here("zones", "boundaries.geojson"))


zones <- zones %>%
  left_join(zone_data)
```

```{r}
# define various color palettes for vis

pnw_contin1 <- pnw_palette(name = "Moth", 
                           n = 5, 
                           type = "continuous")

pnw_contin2 <- pnw_palette(name = "Sunset2", 
                           n = 5, 
                           type = "continuous")

pnw_contin3 <- pnw_palette(name = "Starfish", 
                           n = 5, 
                           type = "continuous")

pnw_contin4 <- pnw_palette(name = "Winter", 
                           n = 5, 
                           type = "continuous")

pnw_contin5 <- pnw_palette(name = "Bay", 
                           n = 5, 
                           type = "continuous")
```

```{r}
# map percentage of low-income households by tract

low_inc_map <- ggplot(zones) +
  geom_sf(aes(fill = pct_low_inc), color = NA)

low_inc_map + 
  theme_void() +
  scale_fill_gradientn(name = "Percent low income \nhouseholds (<$35k)",
                       breaks = my_brks <- seq(0, 1, by = 0.2),
                       labels = paste(my_brks * 100, "%", sep = ""),
                       colors = pnw_contin2)

ggsave(here("images", "a2", "pct_lowinc.png"),
        height = 7,
       width = 7)
```

```{r}
# map percentage of high-income households by tract

high_inc_map <- ggplot(zones) +
  geom_sf(aes(fill = pct_high_inc), color = NA)

high_inc_map +
  theme_void() +
  scale_fill_gradientn(name = "Percent high income \nhouseholds (>$150k)",
                       breaks = my_brks <- seq(0, 1, by = 0.2),
                       labels = paste(my_brks * 100, "%", sep = ""),
                       colors = pnw_contin2)

ggsave(here("images", "a2", "pct_highinc.png"),
        height = 7,
       width = 7)
```

```{r}
# map residential density in MSA

density_map <- ggplot(zones) +
  geom_sf(aes(fill = as.numeric(res_density)), color = NA)

density_map + 
  theme_void() +
  scale_fill_gradientn(name = "Residential density \n(units per square mile)",
                       colors = pnw_contin1)

ggsave(here("images", "a2", "res_dens.png"),
        height = 7,
       width = 7)
```

```{r}
# map racial distribution

race_map <- ggplot(zones) +
  geom_sf(aes(fill = pct_white), color = NA)

race_map +
  theme_void() +
  scale_fill_gradientn(name = "Percent white",
                       breaks = my_brks <- seq(0, 1, by = 0.2),
                       labels = paste(my_brks * 100, "%", sep = ""),
                       colors = pnw_contin3)

ggsave(here("images", "a2", "pct_white.png"),
        height = 7,
       width = 7)
```

```{r}
# generate histogram of vehicle availability

veh_hist <- ggplot(zones) +
  geom_histogram(aes(x = veh_per_hh),
                 fill = "aquamarine",
                 color = "azure4")

veh_hist  +
  theme_classic() +
  scale_x_continuous(name = "Average number of vehicles per household") +
  scale_y_continuous(name = "Number of census tracts")

ggsave(here("images", "a2", "vehicles.png"),
        height = 4.5,
       width = 7)
```

```{r}
# map average commute time by tract

commute_map <- ggplot(zones) +
  geom_sf(aes(fill = avg_commute_time), color = NA)

commute_map +
  theme_void() +
  scale_fill_gradientn(name = "Average commute \ntime (minutes)",
                       colors = pnw_contin5)

ggsave(here("images", "a2", "commute_time.png"),
        height = 7,
       width = 7)
```

```{r}
# generate histogram of commute times by tract

commute_hist <- ggplot(zones) +
  geom_histogram(aes(x = avg_commute_time),
                 fill = "coral2",
                 color = "azure4")

commute_hist +
  theme_classic() +
  scale_x_continuous(name = "Average commute time (minutes)",
                     breaks = seq(0, 135, by = 15)) +
  scale_y_continuous(name = "Number of census tracts")

ggsave(here("images", "a2", "commute_hist.png"),
        height = 4.5,
       width = 7)
```

```{r}
# map number of households with no vehicle access

no_veh_map <- ggplot(zones) +
  geom_sf(aes(fill = hh_no_vehE), color = NA)

no_veh_map + 
  theme_void() +
  scale_fill_gradientn(name = "Number of households \nwithout vehicle access",
                       colors = pnw_contin4)

ggsave(here("images", "a2", "no_veh_map.png"),
        height = 7,
       width = 7)
```

```{r}
# map pct of households with no vehicle access

zones <- zones %>%
  mutate(pct_no_veh = hh_no_vehE/hh_totalE)

no_vh_pct <- ggplot(zones) +
  geom_sf(aes(fill = pct_no_veh), color = NA)

no_vh_pct + 
  theme_void() +
  scale_fill_gradientn(breaks = seq(0, 0.3, by = 0.05), 
                       labels = scales::percent_format(accuracy = 1),
                       name = "Percent of households \nwithout vehicle access",
                       colors = pnw_contin4)

ggsave(here("images", "a2", "no_veh_pct.png"),
        height = 7,
       width = 7)
```

```{r}
# map of employment density

job_dens_map <- ggplot(zones) +
  geom_sf(aes(fill = as.numeric(job_density)), color = NA)

job_dens_map + 
  theme_void() +
  scale_fill_gradientn(name = "Employment density \n(jobs per square mile)",
                       colors = pnw_contin1)

ggsave(here("images", "a2", "job_dens.png"),
        height = 7,
       width = 7)
```

```{r employment stats 1}
# map number of basic jobs by tract

bas_map <- ggplot(zones) +
  geom_sf(aes(fill = basic_emp), color = NA)

bas_hg <- ggplot(zones) +
  geom_histogram(aes(x = basic_emp))

bas_map + 
  theme_void() +
  scale_fill_viridis_c(name = "Number of Jobs \nby Tract (Basic)")

ggsave(here("images", "a2", "basic_jobs.png"),
        height = 7,
       width = 7)
```

```{r employment stats 2}
# map number of retail jobs by tract

ret_map <- ggplot(zones) +
  geom_sf(aes(fill = retail_emp), color = NA)

ret_map + 
  theme_void() +
  scale_fill_gradientn(name = "Number of Jobs \nby Tract (Retail)",
                       colors = pnw_contin2)

ggsave(here("images", "a2", "retail_jobs.png"),
        height = 7,
       width = 7)

ret_hg <- ggplot(zones) +
  geom_histogram(aes(x = retail_emp))
```

```{r employment stats 3}
# map number of service jobs by tract

serv_map <- ggplot(zones) +
  geom_sf(aes(fill = service_emp), color = NA)

serv_map + 
  theme_void() +
  scale_fill_gradientn(name = "Number of Jobs \nby Tract (Service)",
                       colors = pnw_contin5)

ggsave(here("images", "a2", "service_jobs.png"),
        height = 7,
       width = 7)
  
serv_hg <- ggplot(zones) +
  geom_histogram(aes(x = service_emp))
```

```{r}
# map number of commuters who work downtown

commute_dens_map <- ggplot(MSA) +
  geom_sf(aes(fill = as.numeric(commutes)), color = NA) +
  scale_fill_viridis_c(name = "Number of Commuters \nto Downtown by Tract")

commute_dens_map +
  theme_void()

ggsave(here("images", "a2", "dt_commuters.png"),
        height = 7,
       width = 7)
```

```{r the downtown moat}
# map tracts affected by road closures downtown

downtown <- st_union(filter(zones,GEOID == "48453001100"),
                     filter(zones,GEOID == "48453000700"))

moat_map <- ggplot(zones) +
  geom_sf(aes(color = NULL))+
  geom_sf(data = downtown, color = 'red', fill = 'red', lwd = 1)

moat_map +
  theme_void()

ggsave(here("images", "a2", "dt_map.png"),
        height = 7,
       width = 7)
```

```{r}
# generate histograms for employment data by job type

bas_hg
ret_hg
serv_hg
```