---
title: "Generate Combined Skims"
author: "Reuven Herzog and Chris Dsida"
date: "2/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Script Overview

This Rmd doc contains the scripts to generate travel skims.
We will be doing skims for the following modes: car, transit, pedestrian, bicycle.
While we will eventually incorporate transit skims run at multiple times of day, this initial analysis will be limited to a single run at 8:30am.

Our chosen date for the skim is Wednesday, April 6, 2022.

```{r load libraries}
# allocate more memory to Java for using r5r
options(java.parameters = "-Xmx4G")
library(r5r)

library(tidyverse)
library(here)

library(lubridate)
library(sf)
```

```{r load centroids}
centroids <- st_read(here("zones",
                          "centroids.geojson"))
```

## Alternative Network Travel Skims

This section of code chunks generates travel time matrices for driving, walking, biking, and transit in the alternative (No driving allowed in Downtown). Time is calculated for travel between census tract centroids. The maximum allowed travel time is two hours. Code chunk titles describe their function.

```{r setup r5 core}
alt_core <- here("alternative",
                 "networks") %>%
  setup_r5(verbose = FALSE)
```

```{r alt driving skim}
car_skim_alt <- travel_time_matrix(alt_core, 
                     origins = centroids,
                     destinations = centroids,
                     mode = "CAR")
```

```{r alt walking skim}
walk_skim_alt <- travel_time_matrix(alt_core, 
                     origins = centroids,
                     destinations = centroids,
                     mode = "WALK")
```

```{r alt biking skim}
bike_skim_alt <- travel_time_matrix(alt_core, 
                     origins = centroids,
                     destinations = centroids,
                     mode = "BICYCLE")
```

```{r alt transit skim morning rush}
trans_skim_alt_morn <- travel_time_matrix(alt_core, 
                     origins = centroids,
                     destinations = centroids,
                     mode = "TRANSIT",
                     departure_datetime = ymd_hm("2022-04-06 08:30"),
                     breakdown = TRUE)

# Additional code for potentially running skims at other times of day

# trans_skim_alt_eve <- travel_time_matrix(existing_core,
#                       origins = centroids,
#                       destinations = centroids,
#                       mode = "TRANSIT",
#                       departure_datetime = ymd_hm("2022-04-06 17:15"),
#                       breakdown = TRUE)
# 
# 
# trans_skim_alt_midday <- travel_time_matrix(existing_core,
#                       origins = centroids,
#                       destinations = centroids,
#                       mode = "TRANSIT",
#                       departure_datetime = ymd_hm("2022-04-06 12:30"),
#                       breakdown = TRUE)
```

```{r stop existing core}
stop_r5()
```

## Merge and Save Alternative Skims

This section merges the travel time matrices into a single file and saves it to the Github repository. Travel time variable names are changed to correspond to the correct mode.

```{r rename vars and merge}
trans_skim_alt_morn <- trans_skim_alt_morn %>%
  filter(n_rides > 0)

car_skim_alt <- car_skim_alt %>%
  rename(car_time = travel_time) 

trans_skim_alt_morn <- trans_skim_alt_morn %>%
  rename(transit_time = travel_time) 

walk_skim_alt <- walk_skim_alt %>%
  rename(walk_time = travel_time)

bike_skim_alt <- bike_skim_alt %>%
  rename(bike_time = travel_time)

all_skims_alt <- full_join(trans_skim_alt_morn, car_skim_alt) %>%
  full_join(walk_skim_alt) %>%
  full_join(bike_skim_alt)

write_csv(all_skims_alt, here("alternative",
                              "data",
                              "all_skims_alt.csv"))
```

## Existing Network Travel Skims

This section of code chunks generates travel time matrices for driving, walking, biking, and transit under existing conditions. As with the previous section time is calculated for travel between census tract centroids. The maximum allowed travel time is two hours. Code chunk titles describe their function.

```{r setup r5 core}
ex_core <- here("existing",
                 "networks") %>%
  setup_r5(verbose = FALSE)
```

```{r existing driving skim}
car_skim_ex <- travel_time_matrix(ex_core, 
                     origins = centroids,
                     destinations = centroids,
                     mode = "CAR")
```

```{r existing walking skim}
walk_skim_ex <- travel_time_matrix(ex_core, 
                     origins = centroids,
                     destinations = centroids,
                     mode = "WALK")
```

```{r existing biking skim}
bike_skim_ex <- travel_time_matrix(ex_core, 
                     origins = centroids,
                     destinations = centroids,
                     mode = "BICYCLE")
```

```{r existing transit skim morning rush}
trans_skim_ex_morn <- travel_time_matrix(ex_core, 
                     origins = centroids,
                     destinations = centroids,
                     mode = "TRANSIT",
                     departure_datetime = ymd_hm("2022-04-06 08:30"),
                     breakdown = TRUE)
```

```{r stop existing core}
stop_r5()
```

## Merge and Save Existing Skims

This section merges the travel time matrices into a single file and saves it to the Github repository, but for the existing conditions. Travel time variable names are changed to correspond to the correct mode.

```{r rename vars and merge}
trans_skim_ex_morn <- trans_skim_ex_morn %>%
  filter(n_rides > 0)

car_skim_ex <- car_skim_ex %>%
  rename(car_time = travel_time) 

trans_skim_ex_morn <- trans_skim_ex_morn %>%
  rename(transit_time = travel_time) 

walk_skim_ex <- walk_skim_ex %>%
  rename(walk_time = travel_time)

bike_skim_ex <- bike_skim_ex %>%
  rename(bike_time = travel_time)

all_skims_ex <- full_join(trans_skim_ex_morn, car_skim_ex) %>%
  full_join(walk_skim_ex) %>%
  full_join(bike_skim_ex)

write_csv(all_skims_ex, here("existing",
                             "data",
                             "all_skims_ex.csv"))
```

## Convert data for analysis

This section of code sets up the travel time matrices so we can assess whether travel times changed.

```{r loads skims and converts to wide format}
# These lines only necessary if you didn't just run r5r

all_skims_alt <- read.csv(here("alternative",
                              "data",
                              "all_skims_alt.csv"))

all_skims_ex <- read.csv(here("existing",
                             "data",
                             "all_skims_ex.csv"))

# Creates wide format matrix for travel time calculations

# Car, existing conditions

car_mat_ex <- all_skims_ex %>% 
  select(fromId, toId, car_time) %>%
  pivot_wider(names_from = toId, 
              values_from = car_time) %>%
  column_to_rownames("fromId")

# Car, alternative conditions

car_mat_alt <- all_skims_alt %>%
  select(fromId, toId, car_time) %>%
  pivot_wider(names_from = toId, 
              values_from = car_time) %>%
  column_to_rownames("fromId")

# Walking, existing conditions

walk_mat_ex <- all_skims_ex %>%
  select(fromId, toId, walk_time) %>%
  pivot_wider(names_from = toId,
              values_from = walk_time) %>%
  column_to_rownames("fromId")

# Biking, existing conditions

bike_mat_ex <- all_skims_ex %>%
  select(fromId, toId, bike_time) %>%
  pivot_wider(names_from = toId,
              values_from = bike_time) %>%
  column_to_rownames("fromId")

# Transit, existing conditions

trans_mat_ex <- all_skims_ex %>% 
  select(fromId, toId, transit_time) %>%
  pivot_wider(names_from = toId, 
              values_from = transit_time) %>%
  column_to_rownames("fromId")

# These were initially used to compare walking, biking, and transit travel times. They are unchanged between the existing conditions and alternative, so left commented out.

# walk_mat_alt <- all_skims_alt %>%
#   select(fromId, toId, walk_time) %>%
#   pivot_wider(names_from = toId, 
#               values_from = walk_time) %>%
#   column_to_rownames("fromId")
#
# bike_mat_alt <- all_skims_alt %>%
#   select(fromId, toId, bike_time) %>%
#   pivot_wider(names_from = toId, 
#               values_from = bike_time) %>%
#   column_to_rownames("fromId")
#
# trans_mat_alt <- all_skims_alt %>%
#   select(fromId, toId, transit_time) %>%
#   pivot_wider(names_from = toId, 
#               values_from = transit_time) %>%
#   column_to_rownames("fromId")
```

```{r calculate difference in car travel time}
car_diff <- car_mat_ex - car_mat_alt

total_diff <- car_diff %>%
  replace(is.na(.), 0) %>%
  mutate(total_diff = rowSums(across())) %>%
  select(total_diff) %>%
  rownames_to_column("GEOID")

# As noted in the previous chunks, no difference in travel time exists for walking, biking, and transit.

# walk_diff <- walk_mat_ex - walk_mat_alt
# 
# total_diff_walk <- walk_diff %>%
#   replace(is.na(.), 0) %>%
#   mutate(total_diff = rowSums(across())) %>%
#   select(total_diff) %>%
#   rownames_to_column("GEOID")
# 
# bike_diff <- bike_mat_ex - bike_mat_alt
# 
# total_diff_bike <- bike_diff %>%
#   replace(is.na(.), 0) %>%
#   mutate(total_diff = rowSums(across())) %>%
#   select(total_diff) %>%
#   rownames_to_column("GEOID")
#
# trans_diff = trans_mat_ex - trans_mat_alt
# 
# total_diff2 <- trans_diff %>%
#   replace(is.na(.), 0) %>%
#   mutate(total_diff2 = rowSums(across())) %>%
#   select(total_diff2) %>%
#   rownames_to_column("GEOID")
```

```{r calculate travel time difference for a given destination}
# first of two downtown census tracts

existing_max_diff <- car_mat_ex %>%
  rownames_to_column("GEOID") %>%
  select(GEOID, `48453001100`) %>%
  rename(existing = `48453001100`)

alt_max_diff <- car_mat_alt %>%
  rownames_to_column("GEOID") %>%
  select(GEOID, `48453001100`) %>%
  rename(alt = `48453001100`)

# appended to geometry

zones <- here("zones",
              "boundaries.geojson") %>%
  st_read() %>%
  inner_join(existing_max_diff) %>%
  inner_join(alt_max_diff) %>%
  mutate(diff = alt - existing)

# same process for second downtown census tract, with the addition of existing travel times for walking, biking, and public transit. These are used for mapping later

existing_max_diff2 <- car_mat_ex %>%
  rownames_to_column("GEOID") %>%
  select(GEOID, `48453000700`) %>%
  rename(existing = `48453000700`)

alt_max_diff2 <- car_mat_alt %>%
  rownames_to_column("GEOID") %>%
  select(GEOID, `48453000700`) %>%
  rename(alt = `48453000700`)

ex_max_walk2 <- walk_mat_ex %>%
  rownames_to_column("GEOID") %>%
  select(GEOID, `48453000700`) %>%
  rename(walk_ex = `48453000700`)

ex_max_bike2 <- bike_mat_ex %>%
  rownames_to_column("GEOID") %>%
  select(GEOID, `48453000700`) %>%
  rename(bike_ex = `48453000700`)

ex_max_trans2 <- trans_mat_ex %>%
  rownames_to_column("GEOID") %>%
  select(GEOID, `48453000700`) %>%
  rename(trans_ex = `48453000700`)

zones2 <- here("zones",
              "boundaries.geojson") %>%
  st_read() %>%
  inner_join(existing_max_diff2) %>%
  inner_join(alt_max_diff2) %>%
  inner_join(ex_max_walk2) %>%
  inner_join(ex_max_bike2) %>%
  inner_join(ex_max_trans2) %>%
  mutate(diff = alt - existing)
```
## Generate and Save Maps

This section creates a variety of maps used. For driving, the existing travel time is compared to the travel time in the alternative, for both downtown census tracts affected by the proposed road closures. Images are saved to Github repository.

### Existing travel time by car to downtown tract 48453001100

```{r}
library(ggthemes)
library(RColorBrewer)

map_palette <- brewer.pal(5, "YlOrBr")

map_existing <- ggplot(zones) +
  geom_sf(aes(fill = existing),
          color = NA) +
  geom_sf(data = zones[zones$GEOID=="48453001100",],
          fill = "blue",
          color = "blue") +
  scale_fill_gradientn(colours = map_palette,
                       name = "Travel time\n(existing)") +
  theme_void()

map_existing

ggsave("ex_car_times_tract1.png",
       path = here("images",
                   "a4"),
       width = 7.5,
       height = 7.5,
       units = "in")
```

### Alternative travel time by car to downtown tract 48453001100

```{r}
map_alt <- ggplot(zones) +
  geom_sf(aes(fill = alt),
          color = NA) +
  geom_sf(data = zones[zones$GEOID=="48453001100",],
          fill = "blue",
          color = "blue") +
  scale_fill_gradientn(colours = map_palette,
                       name = "Travel time\n(alternative)") +
  theme_void()

map_alt

ggsave("alt_car_times_tract1.png",
       path = here("images",
                   "a4"),
       width = 7.5,
       height = 7.5,
       units = "in")
```

### Difference in car travel time to downtown tract 48453001100

```{r}
map_dif <- ggplot(zones) +
  geom_sf(aes(fill = diff),
          color = NA) +
  geom_sf(data = zones[zones$GEOID=="48453001100",],
          fill = "blue",
          color = "blue") +
  scale_fill_gradientn(colours = map_palette,
                       name = "Difference\nin travel times") +
  theme_void()

map_dif

ggsave("dif_car_times_tract1.png",
       path = here("images",
                   "a4"),
       width = 7.5,
       height = 7.5,
       units = "in")
```

### Existing travel time by car to downtown tract 48453000700

```{r}
map_existing2 <- ggplot(zones2) +
  geom_sf(aes(fill = existing),
          color = NA) +
  geom_sf(data = zones2[zones2$GEOID=="48453000700",],
          fill = "blue",
          color = "blue") +
  scale_fill_gradientn(colours = map_palette,
                       name = "Travel time\n(existing)") +
  theme_void()

map_existing2

ggsave("ex_car_times_tract2.png",
       path = here("images",
                   "a4"),
       width = 7.5,
       height = 7.5,
       units = "in")
```

### Alternative travel time by car to downtown tract 48453000700

```{r}
map_alt2 <- ggplot(zones2) +
  geom_sf(aes(fill = alt),
          color = NA) +
  geom_sf(data = zones[zones$GEOID=="48453000700",],
          fill = "blue",
          color = "blue") +
  scale_fill_gradientn(colours = map_palette,
                       name = "Travel time\n(alternative)") +
  theme_void()

map_alt2

ggsave("alt_car_times_tract2.png",
       path = here("images",
                   "a4"),
       width = 7.5,
       height = 7.5,
       units = "in")
```

### Difference in car travel time to downtown tract 48453000700

```{r}
map_dif2 <- ggplot(zones2) +
  geom_sf(aes(fill = diff),
          color = NA) +
  geom_sf(data = zones[zones$GEOID=="48453000700",],
          fill = "blue",
          color = "blue") +
  scale_fill_gradientn(colours = map_palette,
                       name = "Difference\nin travel times") +
  theme_void()

map_dif2

ggsave("dif_car_times_tract2.png",
       path = here("images",
                   "a4"),
       width = 7.5,
       height = 7.5,
       units = "in")
```

### Existing (and alternative) travel time by walking to tract 48453000700

```{r}
map_ex_walk2 <- ggplot(zones2) +
  geom_sf(aes(fill = walk_ex),
          color = NA) +
  geom_sf(data = zones2[zones2$GEOID=="48453000700",],
          fill = "blue",
          color = "blue") +
  scale_fill_gradientn(colours = map_palette,
                       name = "Travel time\n(existing)") +
  theme_void()

map_ex_walk2

ggsave("ex_walk_times_tract2.png",
       path = here("images",
                   "a4"),
       width = 7.5,
       height = 7.5,
       units = "in")
```

### Existing (and alternative) travel time by biking to tract 48453000700

```{r}
map_ex_bike2 <- ggplot(zones2) +
  geom_sf(aes(fill = bike_ex),
          color = NA) +
  geom_sf(data = zones2[zones2$GEOID=="48453000700",],
          fill = "blue",
          color = "blue") +
  scale_fill_gradientn(colours = map_palette,
                       name = "Travel time\n(existing)") +
  theme_void()

map_ex_bike2

ggsave("ex_bike_times_tract2.png",
       path = here("images",
                   "a4"),
       width = 7.5,
       height = 7.5,
       units = "in")
```

### Existing (and alternative) travel time by transit to tract 48453000700

```{r}
map_ex_trans2 <- ggplot(zones2) +
  geom_sf(aes(fill = trans_ex),
          color = NA) +
  geom_sf(data = zones2[zones2$GEOID=="48453000700",],
          fill = "blue",
          color = "blue") +
  scale_fill_gradientn(colours = map_palette,
                       name = "Travel time\n(existing)") +
  theme_void()

map_ex_trans2

ggsave("ex_trans_times_tract2.png",
       path = here("images",
                   "a4"),
       width = 7.5,
       height = 7.5,
       units = "in")
```