---
title: "Assignment 6"
author: "Chris Dsida and Reuven Herzog"
date: "3/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load packages required for analysis

```{r, results = 'hide', message = FALSE, warning = FALSE}
library(here, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(sf, quietly = TRUE)
library(modelr, quietly = TRUE)
library(jtools, quietly = TRUE)
library(gapminder, quietly = TRUE)
library(ggplot2)
library(GGally)
library(PNWColors)
library(RColorBrewer)
```

## load data for analysis

```{r}
access <- here("existing",
               "data",
               "access_compare.csv") %>%
  read_csv(show_col_types = FALSE)

zones <- here("existing",
               "data",
               "zone_data.csv") %>%
  read_csv(show_col_types = FALSE)
```

## select variables used in model

```{r}
# just copied from resources page; updated with our own vars

model_data <- left_join(zones, access) %>%
  mutate(pct_big_hh = (hh_3personE + hh_4person_plusE) / hh_totalE,
         pct_no_veh = hh_no_vehE / hh_totalE,
         pct_under_18 = pop_under_18E / pop_totalE,
         pct_over_65 = pop_over_65E / pop_totalE,
         pct_basic_emp = basic_emp / total_emp,
         pct_retail_emp = retail_emp / total_emp,
         pct_service_emp = service_emp / total_emp) %>%
  select(GEOID, 
         pct_big_hh, 
         pct_low_inc, 
         pct_high_inc,
         pct_under_18,
         pct_over_65,
         transit_car_ratio, 
         bike_car_ratio, 
         walk_car_ratio, 
         pct_no_veh,
         avg_commute_time,
         res_density,
         job_density,
         job_home_ratio,
         pct_basic_emp,
         pct_service_emp,
         pct_retail_emp)
        
#%>% filter(GEOID != 48453001750)

model_data$GEOID <- as.character(model_data$GEOID)
```

```{r Check Outliers}
#Defining outliers as more than 3 stdevs from the mean

factornames <- colnames(select(model_data,-c("GEOID")))
all_outliers <- data.frame("GEOID" = character())

for (fac in factornames) {
  outlier_test <- select(model_data, c("GEOID", fac))
  outlier_test <- outlier_test %>%
    mutate(not_inf = is.finite(outlier_test[[2]])) %>%
    filter(not_inf == TRUE) %>%
    select(-c("not_inf"))
  
  mn <- mean(outlier_test[[2]])
  stdev <- sd(outlier_test[[2]])
  Tmin <- mn - 3*stdev
  Tmax <- mn + 3*stdev
  #there are no outliers no the left side
  outliers <- filter(outlier_test, outlier_test[[2]] > Tmax)
  
  all_outliers <- all_outliers %>%
    full_join(outliers)  
}

#how do we filter out our outliers?
#I want to say if it is an outlier in at least 2 categories

all_outliers <- all_outliers %>%
  mutate(count_outs = (ncol(all_outliers) - 1) - rowSums(is.na(all_outliers)))

legit_outliers <- all_outliers %>%
  filter(count_outs >= 2)


```

Now we'll filter out our outliers and see if that makes a difference
Four options below, choose which to use by commenting out the other ones

```{r filter outliers}
# filtered_data <- model_data %>%
#   filter(!(GEOID %in% legit_outliers[["GEOID"]]))

# filtered_data <- model_data %>%
#   filter(!(GEOID %in% all_outliers[["GEOID"]]))

# filtered_data <- model_data

filtered_data <- model_data %>%
  filter(GEOID != 48453001750)

```


## exploratory analysis and visualization

```{r}
# this should include plots showing correlations between variables

#just throwing together scatterplots to see what's going on

colnames <- colnames(select(filtered_data,-c("GEOID", "pct_no_veh")))

for (col in colnames) {
  plot <- ggplot(filtered_data, aes(x = filtered_data[[col]], y = pct_no_veh)) +
    geom_point(size = 1, color = "blue") +
    geom_smooth(size = 0.5, lty = "dashed", color = "blue", method = "lm") +
    labs(x = col)
  print(plot)
}
```

```{r}
ggplot(data = filtered_data, aes(x = transit_car_ratio, y = pct_no_veh)) +
  geom_point() +
  labs(x = "Transit-Car Accessibility Ratio", y = "% Without Car Access") +
  geom_smooth(size = 0.5, lty = "dashed", color = "blue", method = "lm") +
  theme_light()

ggsave(here("images", "a6", "transitratio.png"))
```

## estimate regression model

initial thoughts on variables to include:
percent low-/high-income, percent big (3+) households, percent young/old (under 18/over 65), accessibility ratio(s), avg. workers per household, commute time (?), jobs-housing ratio, residential density, ...?
We don't have this in the original variable list but perhaps homeownership percentages?

```{r}
# once vars selected, update with ours

# iterate with model specification, ideally aiming for parsimony


#The Kitchen Sink model
model1 <- lm(pct_no_veh ~ pct_big_hh + pct_low_inc + pct_high_inc + 
                          pct_under_18 + pct_over_65 + transit_car_ratio + 
                          bike_car_ratio + walk_car_ratio + avg_commute_time + 
                          res_density + job_density + job_home_ratio +
                          pct_service_emp + pct_retail_emp, 
             data = filtered_data)

export_summs(model1)
```
```{r}
#only picking the significant vars from model 1

model2 <- lm(pct_no_veh ~ pct_low_inc + transit_car_ratio + walk_car_ratio + 
                          avg_commute_time + res_density, 
             data = filtered_data)

export_summs(model2)
```

```{r}
#only land use variables

model3 <- lm(pct_no_veh ~ transit_car_ratio + 
                          bike_car_ratio + walk_car_ratio + avg_commute_time + 
                          res_density + job_density + job_home_ratio, 
             data = filtered_data)

export_summs(model3)
```

```{r}
#only demographics data
model4 <- lm(pct_no_veh ~ pct_big_hh + pct_low_inc + pct_high_inc + 
                          pct_under_18 + pct_over_65 + pct_service_emp + pct_retail_emp, 
             data = filtered_data)

export_summs(model4)
```

```{r}
#take the best from each category

model5 <- lm(pct_no_veh ~ pct_big_hh + pct_low_inc + pct_high_inc + 
                          pct_under_18 + pct_over_65 + transit_car_ratio + 
                          walk_car_ratio +
                          pct_service_emp + pct_retail_emp, 
             data = filtered_data)

export_summs(model5)


```



Well anyway I have no idea what to make of this.



```{r}
# the no transit model


model6 <- lm(pct_no_veh ~ pct_big_hh + pct_low_inc + pct_high_inc + 
                          pct_under_18 + pct_over_65 + 
                          bike_car_ratio + walk_car_ratio + avg_commute_time + 
                          res_density + job_density + job_home_ratio +
                          pct_service_emp + pct_retail_emp, 
             data = filtered_data)

export_summs(model6)
```


```{r}
#one last stab
model7 <- lm(pct_no_veh ~ pct_big_hh + pct_low_inc + pct_high_inc +
                          pct_under_18 +
                          transit_car_ratio + walk_car_ratio,
             data = filtered_data)

export_summs(model7, model.names = "Last Attempt")
```


```{r}
#we should rename these
export_summs(model1, model2, model3, model4, model5, model6, model7,
             model.names = c("Kitchen Sink","Refined","Just Land Use",
                             "Just Demo","Another Attempt",
                             "W/O Transit Ratio", "Last Attempt"),
             to.file = "html",
             file.name = "full_reg_table.html",
             error_pos = "right")
```

### apply the model to make predictions

```{r}
# once again, just copied from the resources page; needs updates to work for us

alt_access <- here("alternative",
               "data",
               "access_compare.csv") %>%
  read_csv()

alt_zones <- here("alternative",
               "data",
               "zone_data.csv") %>%
  read_csv() %>%
  mutate(pct_big_hh = (hh_3personE + hh_4person_plusE) / hh_totalE,
         pct_no_veh_ex = hh_no_vehE / hh_totalE,
         pct_under_18 = pop_under_18E / pop_totalE) %>%
  left_join(alt_access) %>%
  add_predictions(model = model7) %>%
  mutate(pct_no_veh = ifelse(pred < 0, 0, pred)) %>%
  mutate(hh_no_vehE = round(hh_totalE * pct_no_veh))

alt_zones$GEOID <- as.character(alt_zones$GEOID)

write.csv(alt_zones, 
          here("alternative",
               "data",
               "zone_data.csv"),
          append = FALSE)
```

## predicted accessibility visualization

```{r}
# this section will be for visualization
zone_boundaries <- st_read(here("zones",
                                "boundaries.geojson"))

zones <- zone_boundaries %>%
  left_join(model_data)

pal1 <- brewer.pal(5, "YlOrBr")

pal2 <- pnw_palette(name = "Bay", 
                           n = 5, 
                           type = "continuous")


ggplot(zones) +
  geom_sf(aes(fill = pct_no_veh),
          color = NA) +
  scale_fill_gradientn(colours = pal1,
                       labels = scales::percent) + 
  labs(fill = "% Without Vehicle \nAccess (Existing)") +
  theme_void()

ggsave(here("images", "a6", "act_zones.png"))
```

```{r}
pred_zones <- zone_boundaries %>%
  left_join(alt_zones) %>%
  mutate(pct_no_veh = ifelse(pct_no_veh < 0, 0, pct_no_veh))


ggplot(pred_zones) +
  geom_sf(aes(fill = pct_no_veh),
          color = NA) +
  scale_fill_gradientn(colours = pal1,
                       labels = scales::percent) +
  labs(fill = "% Without Vehicle \nAccess (Predicted)") +
  theme_void()

ggsave(here("images", "a6", "pred_zones.png"))
```

```{r}
pred_zones <- pred_zones %>%
  mutate(change_no_veh = pct_no_veh - pct_no_veh_ex)

ggplot(pred_zones) +
  geom_sf(aes(fill = pct_no_veh),
          color = NA) +
  scale_fill_gradientn(colours = pal2,
                       labels = scales::percent) + 
  labs(fill = "Change in % Without \nVehicle Access") +
  theme_void()

ggsave(here("images", "a6", "change.png"))
```