---
title: "Assignment 6"
author: "Chris Dsida and Reuven Herzog"
date: "3/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load packages required for analysis

```{r, results = 'hide', message = FALSE, warning = FALSE}
library(here, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(sf, quietly = TRUE)
library(modelr, quietly = TRUE)
library(jtools, quietly = TRUE)
```

## load data for analysis

```{r}
access <- here("existing",
               "data",
               "access_compare.csv") %>%
  read_csv(show_col_types = FALSE)

zones <- here("existing",
               "data",
               "zone_data.csv") %>%
  read_csv(show_col_types = FALSE)
```

## select variables used in model

```{r}
# just copied from resources page; update with our own vars

model_data <- left_join(zones, access) %>%
  mutate(pct_big_hh = (hh_3person + hh_4person_plus) / hh_total,
         pct_no_veh = hh_no_veh / hh_total,
         pct_under_18 = pop_under_18 / pop_total,
         pct_over_65 = pop_over_65 / pop_total) %>%
  select(GEOID, pct_big_hh, pct_low_inc, pct_high_inc, ratio, pct_no_veh)
```

## exploratory analysis and visualization

```{r}
# this should include plots showing correlations between variables
```

## estimate regression model

initial thoughts on variables to include:
percent low-/high-income, percent big (3+) households, percent young/old (under 18/over 65), accessibility ratio(s), avg. workers per household, commute time (?), jobs-housing ratio, residential density, ...?
We don't have this in the original variable list but perhaps homeownership percentages?

```{r}
# once vars selected, update with ours

# iterate with model specification, ideally aiming for parsimony

model1 <- lm(pct_no_veh ~ pct_big_hh + pct_low_inc + pct_high_inc + ratio, data = model_data)

export_summs(model1)
```

### apply the model to make predictions

```{r}
# once again, just copied from the resources page; needs updates to work for us

alt_access <- here("alternative",
               "data",
               "access_compare.csv") %>%
  read_csv()

alt_zones <- here("alternative",
               "data",
               "zone_data.csv") %>%
  read_csv() %>%
  left_join(access) %>%
  mutate(pct_big_hh = (hh_3personE + hh_4person_plusE) / total_hhsE,
         pct_lo_inc = inc_quint_1 / total_hhsE,
         pct_hi_inc = inc_quint_5 / total_hhsE,
         pct_no_veh = no_vehE / total_hhsE) %>%
  add_predictions(model = model2) %>%
  mutate(pct_no_veh = pred) %>%
  mutate(no_vehE = round(total_hhsE * pct_no_veh)) 
```

## predicted accessibility visualization

```{r}
# this section will be for visualization
```