---
title: "Generate Skims"
author: "Reuven Herzog"
date: "2/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## script overview

This Rmd doc contains the scripts to generate travel skims.
We will be doing skims for the following modes: car, transit, pedestrian, bicycle.
We also will be running transit skims for 3 times of day:
8:15, 12:30, and 17:15, to correspond to the morning and evening rushes,
plus a mid-day travel time.

Our chosen date for the skim is Wednesday, April 6, 2022.

```{r load libraries}
library(tidyverse)
library(here)

library(lubridate)
library(sf)
```

```{r load centroids}
centroids <- st_read(here("zones",
                          "centroids.geojson"))
```

#Existing network travel skims

```{r setup r5 core}
# allocate more memory to Java for using r5r
options(java.parameters = "-Xmx2G")
library(r5r)

existing_core <- here("existing",
                      "networks") %>%
  setup_r5(verbose = FALSE)
```

```{r driving skim}
car_skim_ext <- travel_time_matrix(existing_core, 
                     origins = centroids,
                     destinations = centroids,
                     mode = "CAR")
```

```{r walking skim, message = FALSE}
walk_skim_ext <- travel_time_matrix(existing_core, 
                     origins = centroids,
                     destinations = centroids,
                     mode = "WALK")
```

```{r biking skim, message = FALSE}
bike_skim_ext <- travel_time_matrix(existing_core, 
                     origins = centroids,
                     destinations = centroids,
                     mode = "BICYCLE")
```

```{r transit skim morning rush, warning = FALSE}
trans_skim_ext_morn <- travel_time_matrix(existing_core, 
                     origins = centroids,
                     destinations = centroids,
                     mode = "TRANSIT",
                     departure_datetime = ymd_hm("2022-04-06 08:15"),
                     breakdown = TRUE)
```

```{r transit skim evening rush, message = FALSE}
trans_skim_ext_eve <- travel_time_matrix(existing_core, 
                     origins = centroids,
                     destinations = centroids,
                     mode = "TRANSIT",
                     departure_datetime = ymd_hm("2022-04-06 17:15"),
                     breakdown = TRUE)
```

```{r transit skim midday, message = FALSE}
trans_skim_ext_midday <- travel_time_matrix(existing_core, 
                     origins = centroids,
                     destinations = centroids,
                     mode = "TRANSIT",
                     departure_datetime = ymd_hm("2022-04-06 12:30"),
                     breakdown = TRUE)
```

```{r stop existing core}
stop_r5()
```

```{r filter transit skims}
tran_skim_ext_morn <- trans_skim_ext_morn %>%
  filter(n_rides > 0)

trans_skim_ext_eve <- trans_skim_ext_eve %>%
  filter(n_rides > 0)

trans_skim_ext_midday <- trans_skim_ext_midday %>%
  filter(n_rides > 0)
```


