---
title: "Assignment 7"
author: "Reuven Herzog"
date: "3/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
library(here)
library(tidyverse)
library(survey)
library(srvyr)
library(naniar)
library(jtools)
library(knitr)
library(huxtable)
library(sf)
library(ggplot2)
library(BAMMtools)
library(scales)
```

## Data Import and Manipulation

First we need to bring in our National Household Travel Survey data.
(Obtained from https://nhts.ornl.gov/)

We already filtered the complete dataset to only Austin data, so it could
fit on our Github repo.
(See adjacent "A7 get only austin data.rmd")

```{r bring in NHTS data}
austin_trips <- read_csv(here(
                        "assignment rmd docs",
                        "a7",
                        "austin_trips.csv"))

austin_hhs <- read_csv(here(
                      "assignment rmd docs",
                      "a7",
                      "austin_hhs.csv"))
```

Code blocks adapted from skills page. This first one creates the variables for HH income quintiles, HH sizes, and vehicle availability.

```{r}
austin_hhs <- austin_hhs %>%
  mutate(inc_quint_ = case_when(HHFAMINC == "01" ~ "1st",
                               HHFAMINC == "02" ~ "1st",
                               HHFAMINC == "03" ~ "1st",
                               HHFAMINC == "04" ~ "2nd",
                               HHFAMINC == "05" ~ "2nd",
                               HHFAMINC == "06" ~ "3rd",
                               HHFAMINC == "07" ~ "4th",
                               HHFAMINC == "08" ~ "5th",
                               HHFAMINC == "09" ~ "5th",
                               HHFAMINC == "10" ~ "5th",
                               HHFAMINC == "11" ~ "5th",
                               TRUE ~ "NA")) %>%
  mutate(size_ = case_when(HHSIZE == 1 ~ "one",
                              HHSIZE == 2 ~ "two",
                              HHSIZE == 3 ~ "three",
                              TRUE ~ "four_plus")) %>%
  mutate(children_ = case_when(HHSIZE > NUMADLT ~ "yes",
                               TRUE ~ "no")) %>%
  mutate(zero_veh_ = (HHVEHCNT == 0)) %>%
  replace_with_na(list(inc_quint_ = "NA")) %>%
  select(HOUSEID, zero_veh_, size_, children_, inc_quint_, WTHHFIN) 
```

The following block creates variables for home-based and work-based trips, then groups them so we know the number of trip types by each household in the survey.

```{r}
trips_by_purpose <- austin_trips %>%
  select(HOUSEID, WHYFROM, WHYTO) %>%
  mutate(home_based = case_when(WHYTO == "01" ~ TRUE,
                                WHYTO == "02" ~ TRUE,
                                WHYFROM == "01" ~ TRUE,
                                WHYFROM == "02" ~ TRUE,
                                TRUE ~ FALSE)) 

trips_by_purpose <- trips_by_purpose %>%
  mutate(work = ifelse(WHYTO == "03" | WHYFROM == "03", TRUE, FALSE)) %>%
  mutate(purpose = case_when(home_based & work ~ "HBW",
                            home_based ~ "HBO",
                            TRUE ~ "NHB")) %>%
  group_by(HOUSEID, purpose)

trips_by_purpose <- trips_by_purpose %>%
  summarize(n = n()) %>%
  pivot_wider(names_from = "purpose", values_from = "n") 
```

Dealing with lots of "NA" entries and replacing them with zeros.
(Note: whereas Carole's tutorial specifically uses a left join to drop trips
outside the Austin MSA, we did this separately already. Nevertheless, a
left join is a perfectly good join to use here.)

```{r}
hh_trips <- left_join(austin_hhs, trips_by_purpose) %>%
  replace_na(list(HBW = 0,
                  HBO = 0,
                  NHB = 0))
```

```{r}
svy_trips <- hh_trips %>%
  as_survey(weights = WTHHFIN)
```

## Trip Production

First HBO model. Same result as the skills page example; only HH size seems to be significant. Added presence of children (members of HH under 18) to the model as well, since we can get that easily in the ACS (in terms of HHs with children)

```{r}
HBO_model1 <- svyglm(HBO ~ zero_veh_ + size_ + inc_quint_ + children_, svy_trips)

export_summs(HBO_model1,
             error_pos = "right",
             error_format = "(p = {p.value})",
             model.names = "Full model") %>%
  quick_rtf(file = here("images",
                       "a7",
                       "HBO Full Model.rtf"))
```

Model attempt #2: now with only HH size. Compared with the first model as well.

```{r}
HBO_model2 <- svyglm(HBO ~ size_, svy_trips)

export_summs(HBO_model1, HBO_model2,
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = c("Full model", "Reduced model")) %>%
  quick_rtf(file = here("images",
                       "a7",
                       "HBO both models.rtf"))
```

This next code block just adds the number of households with children to our zone data and saves it to the repo. No need to run it after the first time.

```{r}
# vars <- c(hh_children = 'B09005_001')
# 
# counties <- c("Travis", 
#               "Bastrop", 
#               "Hays", 
#               "Caldwell", 
#               "Williamson")
# 
# census_data <- get_acs(geography = "tract",
#                        state = "TX",
#                        variables = vars,
#                        county = counties,
#                        output = "wide",
#                        geometry = FALSE) %>%
#   select(GEOID, hh_childrenE)
# 
# census_data$GEOID <- as.numeric(census_data$GEOID)
# 
# existing_zones <- here("existing",
#                        "data",
#                        "zone_data.csv") %>%
#   read_csv() %>%
#   left_join(census_data, by = "GEOID")
# 
# write.csv(existing_zones, here("existing",
#                                "data",
#                                "zone_data.csv"))
```

Apply trip production model to zone data for HBO trips. Because our alternative does not change any of the underlying factors in the trip production model, we do not need to repeat this process and can use the same estimates. 

```{r}
existing_zones <- here("existing",
                       "data",
                       "zone_data.csv") %>%
  read_csv() %>%
  mutate(hbo_prod = hh_totalE * HBO_model2$coefficients["(Intercept)"] +
                    hh_1personE * HBO_model2$coefficients["size_one"] +
                    hh_2personE * HBO_model2$coefficients["size_two"] +
                    hh_3personE * HBO_model2$coefficients["size_three"]) %>%
  mutate(GEOID = as.character(GEOID))
```

Repeating the model process for HBW trips. HH size matters as before, as does the presence of household members under the age of 18. 

```{r}
HBW_model1 <- svyglm(HBW ~ zero_veh_ + size_ + inc_quint_ + children_, svy_trips)

HBW_model2 <- svyglm(HBW ~ size_ + children_, svy_trips)

export_summs(HBW_model1, HBW_model2,
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = c("Full model", "Reduced model")) %>%
  quick_rtf(file = here("images",
                       "a7",
                       "HBW model compare.rtf"))
```

Applying the trip production model to our zone data for HBW trips.

```{r}
existing_zones <- existing_zones %>%
mutate(hbw_prod = hh_totalE * HBW_model2$coefficients["(Intercept)"] +
                  hh_1personE * HBW_model2$coefficients["size_one"] +
                  hh_2personE * HBW_model2$coefficients["size_two"] +
                  hh_3personE * HBW_model2$coefficients["size_three"] +
                  hh_childrenE * HBW_model2$coefficients["children_yes"])
```

Presence of children in the surveyed HHs does not appear to matter here for NHB trips, but income does. Though the coefficient on the 2nd income quintile is statistically insignificant, it is only marginally so.

```{r}
NHB_model1 <- svyglm(NHB ~ zero_veh_ + size_ + inc_quint_ + children_, svy_trips)

NHB_model2 <- svyglm(NHB ~ size_ + inc_quint_, svy_trips)

export_summs(NHB_model1, NHB_model2,
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = c("Full model", "Reduced model")) %>%
  quick_rtf(file = here("images",
                       "a7",
                       "NHB model compare.rtf"))
```

```{r}
existing_zones <- existing_zones %>%
mutate(nhb_prod = hh_totalE * NHB_model2$coefficients["(Intercept)"] +
                  hh_1personE * NHB_model2$coefficients["size_one"] +
                  hh_2personE * NHB_model2$coefficients["size_two"] +
                  hh_3personE * NHB_model2$coefficients["size_three"] +
                  inc_quint_2 * NHB_model2$coefficients["inc_quint_2nd"] +
                  inc_quint_3 * NHB_model2$coefficients["inc_quint_3rd"] +
                  inc_quint_4 * NHB_model2$coefficients["inc_quint_4th"] +
                  inc_quint_5 * NHB_model2$coefficients["inc_quint_5th"])
```

## Trip Attraction

```{r}
existing_zones <- existing_zones %>%
  replace_na(list(basic_emp = 0,
                  retail_emp = 0, 
                  service_emp = 0)) %>%
  mutate(hbo_attr = 0.7 * hh_totalE +
                    0.7 * basic_emp +
                    8.4 * retail_emp +
                    3.5 * service_emp,
         hbw_attr = 1.2 * total_emp,
         nhb_attr = 0.6 * hh_totalE +
                    0.5 * basic_emp +
                    4.7 * retail_emp +
                    1.4 * service_emp)
```

## Balancing Trip Production and Attraction

### HBO Trip Balancing

```{r}
trip_end_summary <- tibble(Purpose = c("HBO"),
                           Productions = c(sum(existing_zones$hbo_prod)),
                           Attractions = c(sum(existing_zones$hbo_attr)),
                           Difference = c(sum(existing_zones$hbo_attr) - 
                                            sum(existing_zones$hbo_prod)))

kable(trip_end_summary, format.args = list(big.mark = ","))
```

```{r}
existing_zones <- existing_zones %>%
  mutate(hbo_attr_bal = hbo_attr * 
           sum(existing_zones$hbo_prod) / sum(existing_zones$hbo_attr))
```

```{r}
trip_end_summary <- tibble(Purpose = c("HBO"),
                           Productions = c(sum(existing_zones$hbo_prod)),
                           Attractions = c(sum(existing_zones$hbo_attr_bal)),
                           Difference = c(sum(existing_zones$hbo_attr_bal) - 
                                            sum(existing_zones$hbo_prod)))

kable(trip_end_summary, format.args = list(big.mark = ","))
```

### HBW Trip Balancing

```{r}
trip_end_summary <- tibble(Purpose = c("HBW"),
                           Productions = c(sum(existing_zones$hbw_prod)),
                           Attractions = c(sum(existing_zones$hbw_attr)),
                           Difference = c(sum(existing_zones$hbw_attr) - 
                                            sum(existing_zones$hbw_prod)))

kable(trip_end_summary, format.args = list(big.mark = ","))
```

```{r}
existing_zones <- existing_zones %>%
  mutate(hbw_attr_bal = hbw_attr * 
           sum(existing_zones$hbw_prod) / sum(existing_zones$hbw_attr))
```

```{r}
trip_end_summary <- tibble(Purpose = c("HBW"),
                           Productions = c(sum(existing_zones$hbw_prod)),
                           Attractions = c(sum(existing_zones$hbw_attr_bal)),
                           Difference = c(sum(existing_zones$hbw_attr_bal) - 
                                            sum(existing_zones$hbw_prod)))

kable(trip_end_summary, format.args = list(big.mark = ","))
```

### NHB Trip Balancing

```{r}
trip_end_summary <- tibble(Purpose = c("NHB"),
                           Productions = c(sum(existing_zones$nhb_prod)),
                           Attractions = c(sum(existing_zones$nhb_attr)),
                           Difference = c(sum(existing_zones$nhb_attr) - 
                                            sum(existing_zones$nhb_prod)))

kable(trip_end_summary, format.args = list(big.mark = ","))
```

```{r}
existing_zones <- existing_zones %>%
  mutate(nhb_attr_bal = nhb_attr * 
           sum(existing_zones$nhb_prod) / sum(existing_zones$nhb_attr))
```

```{r}
trip_end_summary <- tibble(Purpose = c("NHB"),
                           Productions = c(sum(existing_zones$nhb_prod)),
                           Attractions = c(sum(existing_zones$nhb_attr_bal)),
                           Difference = c(sum(existing_zones$nhb_attr_bal) - 
                                            sum(existing_zones$nhb_prod)))

kable(trip_end_summary, format.args = list(big.mark = ","))
```


# Data Viz

Let's make chloropeth maps for trip attractions and productions.
Lol no we're gonna make dot density maps, because tract size varies enough.

First lets turn our data into a spatial object.

```{r}
zone_sf <- read_sf(here("zones",
                        "boundaries.geojson")
                   ) %>%
  mutate(GEOID = as.character(GEOID))

zone_sf <- mutate(zone_sf, area_m2 = unclass(st_area(zone_sf)))

trips_map <- full_join(zone_sf, existing_zones) %>%
  select(GEOID, geometry, area_m2,
         hbo_prod, hbw_prod, nhb_prod,
         hbo_attr, hbw_attr, nhb_attr,
         hbo_attr_bal, hbw_attr_bal, nhb_attr_bal)
```

Let's make the dot density maps

```{r make densities}

hbw_p_pts <- st_sample(trips_map,
                       size = ceiling(trips_map$hbw_prod/1000))

hbo_p_pts <- st_sample(trips_map,
                       size = ceiling(trips_map$hbo_prod/1000))

nhb_p_pts <- st_sample(trips_map,
                       size = ceiling(trips_map$nhb_prod/1000))

hbw_a_pts <- st_sample(trips_map,
                       size = ceiling(trips_map$hbw_attr/1000))

hbo_a_pts <- st_sample(trips_map,
                       size = ceiling(trips_map$hbo_attr/1000))

nhb_a_pts <- st_sample(trips_map,
                       size = ceiling(trips_map$nhb_attr/1000))
```


```{r make density spatial objects}

hbw_p_pts_df <- tibble(trip_type = rep("hbw_prod",length(hbw_p_pts))) %>%
  st_sf(geom = hbw_p_pts)

hbo_p_pts_df <- tibble(trip_type = rep("hbo_prod",length(hbo_p_pts))) %>%
  st_sf(geom = hbo_p_pts)

nhb_p_pts_df <- tibble(trip_type = rep("nhb_prod",length(nhb_p_pts))) %>%
  st_sf(geom = nhb_p_pts)

hbw_a_pts_df <- tibble(trip_type = rep("hbw_attr",length(hbw_a_pts))) %>%
  st_sf(geom = hbw_a_pts)

hbo_a_pts_df <- tibble(trip_type = rep("hbo_attr",length(hbo_a_pts))) %>%
  st_sf(geom = hbo_a_pts)

nhb_a_pts_df <- tibble(trip_type = rep("nhb_attr",length(nhb_a_pts))) %>%
  st_sf(geom = nhb_a_pts)

trips_all_pts <- rbind(hbw_p_pts_df, hbo_p_pts_df, nhb_p_pts_df,
                       hbw_a_pts_df, hbo_a_pts_df, nhb_a_pts_df)

trips_prod_pts <- rbind(hbw_p_pts_df, hbo_p_pts_df, nhb_p_pts_df)
trips_attr_pts <- rbind(hbw_a_pts_df, hbo_a_pts_df, nhb_a_pts_df)

trips_both_hbw <- rbind(hbw_p_pts_df, hbw_a_pts_df)
trips_both_hbo <- rbind(hbo_p_pts_df, hbo_a_pts_df)
trips_both_nhb <- rbind(nhb_p_pts_df, nhb_a_pts_df)

```


##now to actually visualize them

```{r HBW prod}
hbw_dens_map <- ggplot(zone_sf) +
  geom_sf(color = "white") +
  geom_sf(data = hbw_p_pts,
          color = "blue",
          alpha = 0.3,
          size = 0.1) +
  
  theme_void()

hbw_dens_map

```

```{r comparing prod and attr}

hbw_dens_map <- ggplot(zone_sf) +
  geom_sf(color = "white") +
  geom_sf(data = trips_both_hbw,
          aes(color = trip_type),
          alpha = 0.3,
          size = 0.75) +
  scale_color_brewer("Trip Type", palette = "Set1") +
  theme_void() +
  guides(color = guide_legend(override.aes = list(size = 5, alpha = 0.6)))

hbw_dens_map

```
```{r comparing prod and attr}

hbo_dens_map <- ggplot(zone_sf) +
  geom_sf(color = "white") +
  geom_sf(data = trips_both_hbo,
          aes(color = trip_type),
          alpha = 0.3,
          size = 0.75) +
  scale_color_brewer("Trip Type", palette = "Set1") +
  theme_void() +
  guides(color = guide_legend(override.aes = list(size = 5, alpha = 0.6)))

hbo_dens_map

```

```{r comparing prod and attr}

nhb_dens_map <- ggplot(zone_sf) +
  geom_sf(color = "white") +
  geom_sf(data = trips_both_nhb,
          aes(color = trip_type),
          alpha = 0.3,
          size = 0.75) +
  scale_color_brewer("Trip Type", palette = "Set1") +
  theme_void() +
  guides(color = guide_legend(override.aes = list(size = 5, alpha = 0.6)))

nhb_dens_map

```


Well, that should have been predictable.


```{r HBW}
hbw_map <- ggplot(trips_map) +
  geom_sf(aes(fill = hbw_prod), color = NA) +
  scale_fill_viridis_c()

hbw_map
```

```{r HBO}
hbo_map <- ggplot(trips_map) +
  geom_sf(aes(fill = hbo_prod), color = NA) +
  scale_fill_viridis_c()

hbo_map
```


```{r NHB}
nhb_map <- ggplot(trips_map) +
  geom_sf(aes(fill = nhb_prod), color = NA) +
  scale_fill_viridis_c()

nhb_map
```

Hm...not so easy to read. What if we normalized by area?


```{r HBW density}
hbw_map_dens <- ggplot(trips_map) +
  geom_sf(aes(fill = hbw_prod/area_m2), color = NA) +
  scale_fill_viridis_c("Home-based Work Trips\n(Normalized by Tract Area)", option = "B", guide = "none") +
  theme_void()

hbw_map_dens
ggsave(filename = here("images",
                       "a7",
                       "HBW Map.png"))
```
(That's better)

```{r HBW density}
hbo_map_dens <- ggplot(trips_map) +
  geom_sf(aes(fill = hbo_prod/area_m2), color = NA) +
  scale_fill_viridis_c("Home-based Other Trips\n(Normalized by Tract Area)", option = "B", guide = "none") +
  theme_void()

hbo_map_dens
ggsave(filename = here("images",
                       "a7",
                       "HBO Map.png"))
```

```{r HBW density}
nhb_map_dens <- ggplot(trips_map) +
  geom_sf(aes(fill = nhb_prod/area_m2), color = NA) +
  scale_fill_viridis_c("Non-Home Based Trips\n(Normalized by Tract Area)", option = "B", guide = "none") +
  theme_void()

nhb_map_dens
ggsave(filename = here("images",
                       "a7",
                       "NHB Map.png"))
```

##Trip attractions

```{r HBW density}
hbw_a_map_dens <- ggplot(trips_map) +
  geom_sf(aes(fill = hbw_attr/area_m2), color = NA) +
  scale_fill_viridis_c("Home-based Work Trips\n(Normalized by Tract Area)", option = "D", guide = "none") +
  theme_void()

hbw_a_map_dens
ggsave(filename = here("images",
                       "a7",
                       "HBW_A Map.png"))
```


```{r HBW density}
hbo_a_map_dens <- ggplot(trips_map) +
  geom_sf(aes(fill = hbo_attr/area_m2), color = NA) +
  scale_fill_viridis_c("Home-based Other Trips\n(Normalized by Tract Area)", option = "D", guide = "none") +
  theme_void()

hbo_a_map_dens
ggsave(filename = here("images",
                       "a7",
                       "HBO_A Map.png"))
```

```{r HBW density}
nhb_a_map_dens <- ggplot(trips_map) +
  geom_sf(aes(fill = nhb_attr/area_m2), color = NA) +
  scale_fill_viridis_c("Non-Home Based Trips\n(Normalized by Tract Area)", option = "D", guide = "none") +
  theme_void()

nhb_a_map_dens
ggsave(filename = here("images",
                       "a7",
                       "NHB_A Map.png"))
```





##Prod-Attr Comparisonss
What about comparing trip productions and attractions?


```{r HBW prod-attr}
hbw_dif_map <- ggplot(trips_map) +
  geom_sf(aes(fill = hbw_prod - hbw_attr_bal), color = NA) +
  scale_fill_viridis_c("Production/Attraction Difference,\nHome-Based Work Trips",
                       option = "C", limits = c(-20000,NA)) +
  theme_void()

hbw_dif_map
ggsave(filename = here("images",
                       "a7",
                       "HBW Difference Map.png"))
```

```{r HBO prod-attr}
hbo_dif_map <- ggplot(trips_map) +
  geom_sf(aes(fill = hbo_prod - hbo_attr_bal), color = NA) +
  scale_fill_viridis_c("Production/Attraction Difference,\nHome-Based Other Trips",
                       option = "C", limits = c(-100000,NA)) +
  theme_void()

hbo_dif_map
ggsave(filename = here("images",
                       "a7",
                       "HBO Difference Map.png"))
```

```{r HBW prod-attr}
nhb_dif_map <- ggplot(trips_map) +
  geom_sf(aes(fill = nhb_prod - nhb_attr_bal), color = NA) +
  scale_fill_viridis_c("Production/Attraction Difference,\nNon Home-Based",
                       option = "C", limits = c(-50000,NA)) +
  theme_void()

nhb_dif_map
ggsave(filename = here("images",
                       "a7",
                       "NHB Difference Map.png"))
```


What if I take these and normalize _them_ by area?

```{r HBW prod-attr}
hbw_dif_map <- ggplot(trips_map) +
  geom_sf(aes(fill = (hbw_prod - hbw_attr_bal)/area_m2), color = NA) +
  scale_fill_viridis_c("Production/Attraction Difference,\nHome-Based Work Trips",
                       option = "A") +
  theme_void()

hbw_dif_map
ggsave(filename = here("images",
                       "a7",
                       "HBW Difference Map normalized.png"))
```

```{r HBO prod-attr}
hbo_dif_map <- ggplot(trips_map) +
  geom_sf(aes(fill = (hbo_prod - hbo_attr_bal)/area_m2), color = NA) +
  scale_fill_viridis_c("Production/Attraction Difference,\nHome-Based Other Trips",
                       option = "A") +
  theme_void()

hbo_dif_map
ggsave(filename = here("images",
                       "a7",
                       "HBO Difference Map normalized.png"))
```

```{r HBW prod-attr}
nhb_dif_map <- ggplot(trips_map) +
  geom_sf(aes(fill = (nhb_prod - nhb_attr_bal)/area_m2), color = NA) +
  scale_fill_viridis_c("Production/Attraction Difference,\nNon Home-Based",
                       option = "A") +
  theme_void()

nhb_dif_map
ggsave(filename = here("images",
                       "a7",
                       "NHB Difference Map normalized.png"))
```


Now I'm curious about if we separate into prod-surplus and attr-surplus

Production surplus
```{r HBW prod-attr}
hbw_dif_map_prod <- ggplot(trips_map) +
  geom_sf(aes(fill = hbw_prod - hbw_attr_bal), color = NA) +
  scale_fill_viridis_c("Production/Attraction Difference,\nHome-Based Work Trips",
                       option = "C", limits = c(0,NA)) +
  theme_void()

hbw_dif_map_prod
ggsave(filename = here("images",
                       "a7",
                       "HBW Difference Map - prod surplus.png"))
```

```{r HBO prod-attr}
hbo_dif_map_prod <- ggplot(trips_map) +
  geom_sf(aes(fill = hbo_prod - hbo_attr_bal), color = NA) +
  scale_fill_viridis_c("Production/Attraction Difference,\nHome-Based Other Trips",
                       option = "C", limits = c(0,NA)) +
  theme_void()

hbo_dif_map_prod
ggsave(filename = here("images",
                       "a7",
                       "HBO Difference Map - prod surplus.png"))
```

```{r HBW prod-attr}
nhb_dif_map_prod <- ggplot(trips_map) +
  geom_sf(aes(fill = nhb_prod - nhb_attr_bal), color = NA) +
  scale_fill_viridis_c("Production/Attraction Difference,\nNon Home-Based",
                       option = "C", limits = c(0,NA)) +
  theme_void()

nhb_dif_map_prod
ggsave(filename = here("images",
                       "a7",
                       "NHB Difference Map - prod surplus.png"))
```


And now attraction-surplus
```{r HBW prod-attr}
hbw_dif_map_attr <- ggplot(trips_map) +
  geom_sf(aes(fill = hbw_prod - hbw_attr_bal), color = NA) +
  scale_fill_viridis_c("Production/Attraction Difference,\nHome-Based Work Trips",
                       option = "C", limits = c(NA,0)) +
  theme_void()

hbw_dif_map_attr
ggsave(filename = here("images",
                       "a7",
                       "HBW Difference Map.png"))
```

```{r HBO prod-attr}
hbo_dif_map_attr <- ggplot(trips_map) +
  geom_sf(aes(fill = hbo_prod - hbo_attr_bal), color = NA) +
  scale_fill_viridis_c("Production/Attraction Difference,\nHome-Based Other Trips",
                       option = "C", limits = c(NA,0)) +
  theme_void()

hbo_dif_map_attr
ggsave(filename = here("images",
                       "a7",
                       "HBO Difference Map.png"))
```

```{r HBW prod-attr}
nhb_dif_map_attr <- ggplot(trips_map) +
  geom_sf(aes(fill = nhb_prod - nhb_attr_bal), color = NA) +
  scale_fill_viridis_c("Production/Attraction Difference,\nNon Home-Based",
                       option = "C", limits = c(NA,0)) +
  theme_void()

nhb_dif_map_attr
ggsave(filename = here("images",
                       "a7",
                       "NHB Difference Map.png"))
```

Other visualization:
Share of each type of trip by zone

## Ratio (not absolute) of Prod/Attr

Let's first generate the ratio, and then bin them

```{r generate ratios and bins}

trips_map <- trips_map %>%
  mutate(hbw_rat = hbw_prod/hbw_attr_bal,
         hbo_rat = hbo_prod/hbo_attr_bal,
         nhb_rat = nhb_prod/nhb_attr_bal)


hbw_breaks <- getJenksBreaks(var = trips_map$hbw_rat, k = 7)
hbo_breaks <- getJenksBreaks(var = trips_map$hbo_rat, k = 7)
nhb_breaks <- getJenksBreaks(var = trips_map$nhb_rat, k = 7)

trips_map <- trips_map %>%
  mutate(hbw_rat_bin_jenks = case_when(
            hbw_rat > hbw_breaks[7] ~ paste("< ", round(max(hbw_rat),2)),
            hbw_rat > hbw_breaks[6] ~ paste("< ", round(hbw_breaks[7],2)),
            hbw_rat > hbw_breaks[5] ~ paste("< ", round(hbw_breaks[6],2)),
            hbw_rat > hbw_breaks[4] ~ paste("< ", round(hbw_breaks[5],2)),
            hbw_rat > hbw_breaks[3] ~ paste("< ", round(hbw_breaks[4],2)),
            hbw_rat > hbw_breaks[2] ~ paste("< ", round(hbw_breaks[3],2)),
            hbw_rat > hbw_breaks[1] ~ paste("< ", round(hbw_breaks[2],2)),
            TRUE ~ paste("< ", round(hbw_breaks[1],2))),
         hbo_rat_bin_jenks = case_when(
            hbo_rat > hbo_breaks[7] ~ paste("< ", round(max(hbo_rat),2)),
            hbo_rat > hbo_breaks[6] ~ paste("< ", round(hbo_breaks[7],2)),
            hbo_rat > hbo_breaks[5] ~ paste("< ", round(hbo_breaks[6],2)),
            hbo_rat > hbo_breaks[4] ~ paste("< ", round(hbo_breaks[5],2)),
            hbo_rat > hbo_breaks[3] ~ paste("< ", round(hbo_breaks[4],2)),
            hbo_rat > hbo_breaks[2] ~ paste("< ", round(hbo_breaks[3],2)),
            hbo_rat > hbo_breaks[1] ~ paste("< ", round(hbo_breaks[2],2)),
            TRUE ~ paste("< ", round(hbo_breaks[1],2))),
         nhb_rat_bin_jenks = case_when(
            nhb_rat > nhb_breaks[7] ~ paste("< ", round(max(nhb_rat),2)),
            nhb_rat > nhb_breaks[6] ~ paste("< ", round(nhb_breaks[7],2)),
            nhb_rat > nhb_breaks[5] ~ paste("< ", round(nhb_breaks[6],2)),
            nhb_rat > nhb_breaks[4] ~ paste("< ", round(nhb_breaks[5],2)),
            nhb_rat > nhb_breaks[3] ~ paste("< ", round(nhb_breaks[4],2)),
            nhb_rat > nhb_breaks[2] ~ paste("< ", round(nhb_breaks[3],2)),
            nhb_rat > nhb_breaks[1] ~ paste("< ", round(nhb_breaks[2],2)),
            TRUE ~ paste("< ", round(nhb_breaks[1],2))))



```



```{r HBW prod/attr}
hbw_rat_map <- ggplot(trips_map) +
  geom_sf(aes(fill = hbw_rat_bin_jenks), color = NA) +
  scale_fill_viridis_d("Production/Attraction Ratio,\nHome-Based Work Trips",
                       labels = c("< 0", "< 2.11", "< 5.32", "< 11.75", "< 24.35", "< 52.77", "< 114.01"),
                       option = "A") +
  theme_void()

hbw_rat_map
ggsave(filename = here("images",
                       "a7",
                       "HBW Ratio Map.png"))
```


```{r hbo prod/attr}
hbo_rat_map <- ggplot(trips_map) +
  geom_sf(aes(fill = hbo_rat_bin_jenks), color = NA) +
  scale_fill_viridis_d("Production/Attraction Ratio,\nHome-Based Work Trips",
                       #labels = c("< 0", "< 2.11", "< 5.32", "< 11.75", "< 24.35", "< 52.77", "< 114.01"),
                       option = "A") +
  theme_void()

hbo_rat_map
ggsave(filename = here("images",
                       "a7",
                       "hbo Ratio Map.png"))
```


```{r nhb prod/attr}
nhb_rat_map <- ggplot(trips_map) +
  geom_sf(aes(fill = nhb_rat_bin_jenks), color = NA) +
  scale_fill_viridis_d("Production/Attraction Ratio,\nHome-Based Work Trips",
                       labels = c("< 0", "< 2.11", "< 5.32", "< 11.75", "< 24.35", "< 52.77", "< 114.01"),
                       option = "A") +
  theme_void()

nhb_rat_map
ggsave(filename = here("images",
                       "a7",
                       "nhb Ratio Map.png"))
```